---
title: "Action Analysis"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("R/data.R")
library(mgcv)
library(broom)
library(ggplot2)
```

# Data

A GAM of Furrycat's data shows a strong relationship between action and both dexterity and intellect.

```{r echo=TRUE, message=FALSE, warning=FALSE}
model <- gam(
  formula = action ~
    s(hardiness) +
    s(fortitude) + 
    s(dexterity) + 
    s(endurance) +
    s(intellect) + 
    s(cleverness) + 
    s(courage) + 
    s(dependability) +
    s(power) +
    s(fierceness) +
    armor,
  family = gaussian(),
  data = normalized_df
)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
summary(model)
```

And the GBM shows linear relationships.

```{r echo=TRUE, message=FALSE, warning=FALSE}
plot(model, select = 3)
plot(model, select = 5)
```


The linear model shows high correlation and low residuals.


```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
model <- lm(action ~ dexterity + intellect, data = normalized_df)
summary(model)
```

And looks like this.


```{r echo=FALSE, message=FALSE, warning=FALSE}
normalized_df$predicted_action <- predict.lm(model, newdata = normalized_df)

ggplot(normalized_df, aes(x = predicted_action, y = action)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  ggtitle("Linear Model Predicted Action vs Actual")

```

The residuals suggest that some of the error may be due to rounding.

```{r echo=FALSE, message=FALSE, warning=FALSE}
res <- resid(model)
qqnorm(res)
qqline(res)
```

# Conclusion

Action is roughly captured by the following formula, $action \approx 42 + 15 * dexterity + 3 * intellect$. The final value is rounded.
