---
title: "Attribute Analysis"
output: html_document
params:
  attribute_name: "default_attribute"
  major_variable_name: "default_major_variable"
  minor_variable_name: "default_minor_variable"
---


# `r params$attribute_name` Analysis

The GBM of Furrycat's data shows a strong relationship between `r params$attribute_name` and both `r params$major_variable_name` and `r params$minor_variable_name`.

```{r}
library(magrittr)
library(dplyr)
library(readr)
library(ggplot2)
library(gbm)

# Some of these are obvious typos -- fix later
bad_data <- c("dta2275t", "r8tfng1v",              # weird hardiness
              "er9nq4et", "38u9raer", "6ko7k4ql",  # weird action/dex
              "6td1segp", "6f9lmdbu",              # bad speed
              "frfksnne",                          # Maybe bad to-hit but interesting...
              "kjvtv7d7",                          # has fortitude < 500 AND light armor?
              "4231uemm",                          # trash mob in piket skin (min cl 25) that throws off CL calculations
              "at8d24tc",                          # rancor skin (min cl 35) also throwing off cl calculations
              "8mgpmeev",                          # thune skin also throwing off cl
              "v22scdcg",                          # fambaa
              "ceh4m90v",                          # kimogila
              "mm8gn9sj"                           # kimogila
)
creatures <- read_csv("data/clean/furrycat/creatures.csv") %>%
  filter(!serial %in% bad_data)
templates <- read_csv("data/clean/furrycat/templates.csv")
combined_data <- creatures %>% inner_join(templates, by = join_by(template_id == serial))
normalized_df <- combined_data %>%
  mutate_at(
    vars("kinetic.effective.x",
         "kinetic.special.x",
         "energy.effective.x",
         "energy.special.x",
         "blast.effective.x",
         "blast.special.x",
         "heat.effective.x",
         "heat.special.x",
         "cold.effective.x",
         "cold.special.x",
         "electricity.effective.x",
         "electricity.special.x",
         "acid.effective.x",
         "acid.special.x",
         "stun.effective.x",
         "stun.special.x"),
    ~ ifelse(is.na(.), 0, .)
  )

experiments <- read_csv("data/clean/furrycat/experiments.csv")
samples <- read_csv("data/clean/furrycat/samples.csv")
```

```{r}
1+1
```

The linear model shows high correlation and low residuals.


```{r}
2 + 2
```

And looks like this.


```{r}


```


# Conclusion

Health is roughly captured by the following formula, `health ~ 35 + hardiness * 15 + dexterity * 3`.
