---
title: "Creature Level Analysis"
output: html_document
---

# Data

The GBM of Furrycat's data shows a strong relationship between creature level and, well, nearly everything.


## The Problems with Creature Level

There are several big problems with attempting to reverse-engineer the creature level algorithm:

- Bio-Engineers created pets for market demand, so the data set is heavily biased, preferring kinetic resists, optimizing for health, action, and mind -- in that order -- etc.
- So-called "vulnerability stacking" complicates things considerably and it was, at best, an oversight; and at worst, a bug

# The Creature Level Algorithm is Probably Multiplicative

```{r message=FALSE, warning=FALSE, include=FALSE}
library(magrittr)
library(dplyr)
library(readr)
library(ggplot2)
library(gbm)

# Some of these are obvious typos -- fix later
bad_data <- c("dta2275t", "r8tfng1v",              # weird hardiness
              "er9nq4et", "38u9raer", "6ko7k4ql",  # weird action/dex
              "6td1segp", "6f9lmdbu",              # bad speed
              "frfksnne",                          # Maybe bad to-hit but interesting...
              "kjvtv7d7",                          # has fortitude < 500 AND light armor?
              "4231uemm",                          # trash mob in piket skin (min cl 25) that throws off CL calculations
              "at8d24tc",                          # rancor skin (min cl 35) also throwing off cl calculations
              "8mgpmeev",                          # thune skin also throwing off cl
              "v22scdcg",                          # fambaa
              "ceh4m90v",                          # kimogila
              "mm8gn9sj"                           # kimogila
)
creatures <- read_csv("data/clean/furrycat/creature.csv") %>%
  filter(!creature_id %in% bad_data)
final_exp <- read_csv("data/clean/furrycat/final_experiment.csv")
combined_data <- inner_join(creatures, final_exp, by = c("final_experimentation_id" = "serial"))
normalized_df <- combined_data %>%
  mutate_at(
    vars("kinetic.x", "energy.x", "blast.x", "heat.x", "cold.x", "electricity.x", "acid.x", "stun.x"),
    ~ ifelse(is.na(.), 0, .)
  )
```


Let's consider "non-vulnerable" creatures for the moment -- this is simply to cut out "vuln stacking" as a complication.

```{r echo=TRUE, message=FALSE, warning=FALSE}
non_vuln <- normalized_df %>% 
  filter(kinetic.x >= 0, 
         energy.x >= 0, 
         blast.x >= 0, 
         heat.x >= 0, 
         cold.x >= 0, 
         electricity.x >= 0, 
         acid.x >= 0, 
         stun.x >= 0)
```

This very interestingly fits a Gamma distribution!

```{r echo=TRUE, message=FALSE, warning=FALSE}
gamma_model <- glm(creature_level ~ hardiness + fortitude + dexterity + endurance + intellect + cleverness + courage + dependability + fierceness + power + armor.x + kinetic.x + energy.x + blast.x + heat.x + cold.x +electricity.x + acid.x + stun.x, data = non_vuln, family = Gamma())

residuals <- resid(gamma_model)
qqnorm(residuals)
qqline(residuals)
```

And the Shapiro-Wilks test for this is:

```{r echo=TRUE, message=FALSE, warning=FALSE}
shapiro.test(residuals)
```

Which is very suggestive that the residuals are normally distributed -- meaning, the variations in creature level are very likely simply due to random variations in assembly, experimentation, and so on.

# Further

```{r echo=TRUE, message=FALSE, warning=FALSE}
gbm_model <- gbm(
  creature_level ~ hardiness + fortitude + dexterity + endurance + intellect + cleverness + courage + dependability + fierceness + power + armor.x + kinetic.x + energy.x + blast.x + heat.x + cold.x +  electricity.x + acid.x + stun.x,
  data = normalized_df,
  distribution = "gaussian",
  n.trees = 500,
  interaction.depth = 3,
  n.minobsinnode = 10,
  bag.fraction = 0.3,
  cv.folds = 10
)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
summary(gbm_model, plotit = FALSE)
```

The GBM shows several relationships, however, it is likely that not all are valid given biases in the data set.

```{r echo=TRUE, message=FALSE, warning=FALSE}
plot.gbm(gbm_model, i.var = "intellect", n.trees = 500)
```
