---
title: "Creature Level Analysis"
output: html_document
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("R/data.R")
library(mgcv)
library(broom)
library(ggplot2)
library(lmtest)

no_armor_df <- normalized_df %>% filter(armor == 0)
armor_df <- normalized_df %>% filter(armor == 1 & fortitude >= 500)
```

# Data

The GAM model of creature level shows relative influence to a few attributes.

```{r echo=TRUE, message=FALSE, warning=FALSE}

model.gam.everything <- gam(
  level ~
    s(hardiness) +
    s(fortitude) +
    s(dexterity) +
    s(endurance) +
    s(intellect) +
    s(cleverness) +
    s(dependability) +
    s(courage) +
    s(fierceness) +
    s(power) +
    s(kinetic) +
    s(energy)  +
    s(blast) +
    s(heat) +
    s(cold) +
    s(electricity) +
    s(acid) +
    s(stun),
  data = normalized_df,
  family = gaussian()
)

summary(model.gam.everything)

```
Specifically, the following attributes:

- hardiness
- fortitude 
- dexterity
- intellect
- cleverness
- power
- kinetic
- energy
- cold

The appearance of `kinetic`, `energy`, `cold` and no other resists is especially strange.

Also interesting is that the following attributes have `1.00` degrees of freedom (so their influence is essentially flat):

- endurance
- fierceness
- blast
- heat
- electricity

## Feature Engineering

We are going to make some assumptions about the model based on our domain knowledge
of the game, and to avoid over-fitting.

1. Hardiness, Dexterity, and Intellect have the same weights against level

It's possible that the additive models associate higher weights to hardiness due to
its correlation with fortitude. However, for now, we'll assume that each attribute is more 
or less equal.

2. Kinetic and Energy resists have the same weights against level

We're combining these two because they are uniquely capped at 60%.

3. The other resists have the same weight against level

This is due to domain knowledge -- "vuln stacking" creatures caused a measurable drop
in creature level. It's possible the special emphasis given to `cold` is due to over-fitting or something unique about furrycat's data.

To mediate these concerns, we can create the following synthetic attributes

1. `average_hdi` -- is the average of hardiness, dexterity, and intellect. Taking the mean of these attributes and training the GAM on this synthetic feature will force it not to over-fit on any of hardiness, dexterity, and intellect.

2. `kinen` -- mean of kinetic and energy, for the same reasons.

3. `nonkinen` -- mean of cold, heat, electricity, acid, and stun.

With that, this is the final model:

```{r echo=TRUE, message=FALSE, warning=FALSE}
model.gam <- gam(
  level ~
    s(average_hdi) +
    s(fortitude) +
    s(cleverness) +
    s(power) +
    s(kinen) +
    s(nonkinen),
  data = normalized_df
)
summary(model.gam)
```


## Armored Creature Levels

Analysis of the data shows extreme an extreme segmentation point $fortitude = 500$.

In fact, the $fortitude >= 500$ data set is particularly well-behaved.

```{r}
model.gam.armor <- gam(
  level ~
    s(average_hdi) +
    s(fortitude) +
    s(cleverness) +
    s(power) +
    s(kinen) +
    s(nonkinen),
  data = armor_df
)
summary(model.gam.armor)
```

In fact, the relative degrees of freedom show that several of these parameters are already close to linear.

```{r}
linear.fit.level.armor <- lm(
  level ~
    average_hdi +
    fortitude +
    cleverness +
    power +
    kinen +
    nonkinen,
  data = armor_df
)
summary(linear.fit.level.armor)
```

This simple linear model behaves remarkably well.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(armor_df, aes(x = predict(linear.fit.level.armor, newdata = armor_df), y = level)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Predicted Level vs Actual", x = "predicted_level", y = "level")
```

And analysis of the residuals are highly promising.

```{r echo=FALSE, message=FALSE, warning=FALSE}
rs.model.level.armor <- resid(linear.fit.level.armor)
qqnorm(rs.model.level.armor)
qqline(rs.model.level.armor)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
shapiro.test(rs.model.level.armor)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
bptest(linear.fit.level.armor)
```

To summarize, this simple linear model for the "armored" data set (i.e. creatures that have armor):

1. The residuals are +/- 4 levels 
2. The residuals have median 0.19 levels.
2. The residuals are normally distributed
3. The residuals are [homoscedastic](https://en.wikipedia.org/wiki/Homoscedasticity_and_heteroscedasticity)

Taken together, this may imply that the residuals are *due to randomness within the crafting system itself*, and may not be due to missing variables or unknown non-linear relationships.

## Unarmored Creature Levels

The same analysis for unarmored creatures does not look as promising.

```{r echo=TRUE, message=FALSE, warning=FALSE}
linear.fit.level.noarmor <- lm(
  level ~
    average_hdi +
    fortitude +
    cleverness +
    power +
    kinen +
    nonkinen,
  data = no_armor_df
)
summary(linear.fit.level.noarmor)
```

Which shows a statistically decent but clearly heteroscedastic fit.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(no_armor_df, aes(x = predict(linear.fit.level.noarmor, newdata = no_armor_df), y = level)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Predicted Level vs Actual", x = "predicted_level", y = "level")
```

The residuals are clearly not normal.

```{r echo=FALSE, message=FALSE, warning=FALSE}
rs.model.level.noarmor <- resid(linear.fit.level.noarmor)
qqnorm(rs.model.level.noarmor)
qqline(rs.model.level.noarmor)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
shapiro.test(rs.model.level.noarmor)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
bptest(linear.fit.level.noarmor)
```

### Summary and areas for future investigation

1. Using segment analysis, there is statistical evidence that armor is a change point in the model for creature level
2. An ordinary linear regression fits armored creature levels remarkably well
3. Unarmored creature levels are not so simple...

For unarmored creature level, a few ideas:

1. It could be that lower level are more sensitive to rounding, giving the heteroscedastic results
2. Creature level may be calculated from "final" attributes (such as health, action, mind), which themselves are rounded, which further causes sensitivity at lower creature levels and attributes, where rounding is proportionally more significant
3. There may be more change points in the unarmored data set

### The Strange Case of Fortitude

The influence of fortitude shows a slight _negative_ influence below $fortitude = 500$, and a positive influence above.

See the influence plot,

```{r echo=FALSE, message=FALSE, warning=FALSE}
plot(model.gam, select = 2)
```

This is reflected in the above linear model, where fortitude's slope in the unarmored data is actually negative!

Recall the result of the linear, unarmored model:

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(linear.fit.level.noarmor)
```

Can this be true? Other influence models, such as GBMs, also show the same thing. Still, it is not an intuitive finding. However, note that:

1. It would explain the existence of high-fortitude, unarmored CL 10s with lots of vulnerabilities -- maximizing fortitude to be as close to 499 as possible would _minimize_ its contribution to creature level!
2. It would explain the question of how many levels armor adds -- some players said 10, others said 6. In this view, obtaining armor does not strictly add a constant creature level; it rolls you into a different segment of the linear model, where the influence of fortitude becomes sharply positive
3. It would also explain why change point detection was so adamant about a change point existing at $fortitude = 500$ -- it's not that armor adds a constant level, it's that parameters of the level calculation change entirely!

I'm unsure if this could be accurate, but it is compelling nonetheless.

