---
title: "Creature Level Analysis"
output: html_document
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("R/data.R")
library(mgcv)
library(broom)
library(ggplot2)
library(lmtest)

no_armor_df <- normalized_df %>% filter(armor == 0)
armor_df <- normalized_df %>% filter(armor == 1 & fortitude >= 500)
```

# Data

The GAM model of creature level shows relative influence to a few attributes.

```{r echo=TRUE, message=FALSE, warning=FALSE}

model.gam.everything <- gam(
  level ~
    s(hardiness) +
    s(fortitude) +
    s(dexterity) +
    s(endurance) +
    s(intellect) +
    s(cleverness) +
    s(dependability) +
    s(courage) +
    s(fierceness) +
    s(power) +
    s(kinetic) +
    s(energy)  +
    s(blast) +
    s(heat) +
    s(cold) +
    s(electricity) +
    s(acid) +
    s(stun),
  data = normalized_df,
  family = gaussian()
)

summary(model.gam.everything)

```
Specifically, the following attributes:

- hardiness
- fortitude 
- dexterity
- intellect
- cleverness
- power
- kinetic
- energy
- cold

The appearance of `kinetic`, `energy`, `cold` and no other resists is especially strange.

Also interesting is that the following attributes have `1.00` degrees of freedom (so their influence is essentially flat):

- endurance
- fierceness
- blast
- heat
- electricity

## Feature Engineering

We are going to make some assumptions about the model based on our domain knowledge
of the game, and to avoid over-fitting.

1. Hardiness, Dexterity, and Intellect have the same weights against level

It's possible that the additive models associate higher weights to hardiness due to
its correlation with fortitude. However, for now, we'll assume that each attribute is more 
or less equal.

2. Kinetic and Energy resists have the same weights against level

We're combining these two because they are uniquely capped at 60%.

3. The other resists have the same weight against level

This is due to domain knowledge -- "vuln stacking" creatures caused a measurable drop
in creature level. It's possible the special emphasis given to `cold` is due to over-fitting or something unique about furrycat's data.

To mediate these concerns, we can create the following synthetic attributes

1. `average_hdi` -- is the average of hardiness, dexterity, and intellect. Taking the mean of these attributes and training the GAM on this synthetic feature will force it not to over-fit on any of hardiness, dexterity, and intellect.

2. `kinen` -- mean of kinetic and energy, for the same reasons.

3. `nonkinen` -- mean of cold, heat, electricity, acid, and stun.

With that, this is the final model:

```{r echo=TRUE, message=FALSE, warning=FALSE}
model.gam <- gam(
  level ~
    s(average_hdi) +
    s(fortitude) +
    s(cleverness) +
    s(power) +
    s(kinen) +
    s(nonkinen),
  data = normalized_df
)
summary(model.gam)
```


## Armored Creature Levels

Analysis of the data shows extreme an extreme segmentation point $fortitude = 500$.

In fact, the $fortitude >= 500$ data set is particularly well-behaved.

```{r}
model.gam.armor <- gam(
  level ~
    s(average_hdi) +
    s(fortitude) +
    s(cleverness) +
    s(power) +
    s(kinen) +
    s(nonkinen),
  data = armor_df
)
summary(model.gam.armor)
```

In fact, the relative degrees of freedom show that several of these parameters are already close to linear.

```{r}
linear.fit.level.armor <- lm(
  level ~
    average_hdi +
    fortitude +
    cleverness +
    power +
    kinen +
    nonkinen,
  data = armor_df
)
summary(linear.fit.level.armor)
```

This simple linear model behaves remarkably well.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(armor_df, aes(x = predict(linear.fit.level.armor, newdata = armor_df), y = level)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Predicted Level vs Actual", x = "predicted_level", y = "level")
```

And analysis of the residuals are highly promising.

```{r echo=FALSE, message=FALSE, warning=FALSE}
rs.model.level.armor <- resid(linear.fit.level.armor)
qqnorm(rs.model.level.armor)
qqline(rs.model.level.armor)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
shapiro.test(rs.model.level.armor)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
bptest(linear.fit.level.armor)
```

To summarize, this simple linear model for the "armored" data set (i.e. creatures that have armor):

1. The residuals are +/- 4 levels 
2. The residuals have median 0.19 levels.
2. The residuals are normally distributed
3. The residuals are [homoscedastic](https://en.wikipedia.org/wiki/Homoscedasticity_and_heteroscedasticity)

Taken together, this may imply that the residuals are *due to randomness within the crafting system itself*, and may not be due to missing variables or unknown non-linear relationships.

## Unarmored Creature Levels

The same analysis for unarmored creatures does not look as promising.

```{r echo=TRUE, message=FALSE, warning=FALSE}
linear.fit.level.noarmor <- lm(
  level ~
    average_hdi +
    fortitude +
    cleverness +
    power +
    kinen +
    nonkinen,
  data = no_armor_df
)
summary(linear.fit.level.noarmor)
```

Which shows a statistically decent but clearly heteroscedastic fit.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(no_armor_df, aes(x = predict(linear.fit.level.noarmor, newdata = no_armor_df), y = level)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Predicted Level vs Actual", x = "predicted_level", y = "level")
```

The residuals are clearly not normal.

```{r echo=FALSE, message=FALSE, warning=FALSE}
rs.model.level.noarmor <- resid(linear.fit.level.noarmor)
qqnorm(rs.model.level.noarmor)
qqline(rs.model.level.noarmor)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
shapiro.test(rs.model.level.noarmor)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
bptest(linear.fit.level.noarmor)
```

### Summary and areas for future investigation

1. Using segment analysis, there is statistical evidence that armor is a change point in the model for creature level
2. An ordinary linear regression fits armored creature levels remarkably well
3. Unarmored creature levels are not so simple...

For unarmored creature level, a few ideas:

1. It could be that lower level are more sensitive to rounding, giving the heteroscedastic results
2. Creature level may be calculated from "final" attributes (such as health, action, mind), which themselves are rounded, which further causes sensitivity at lower creature levels and attributes, where rounding is proportionally more significant
3. There may be more change points in the unarmored data set

### The Strange Case of Fortitude

The influence of fortitude shows a slight _negative_ influence below $fortitude = 500$, and a positive influence above.

See the influence plot,

```{r echo=FALSE, message=FALSE, warning=FALSE}
plot(model.gam, select = 2)
```

This is reflected in the above linear model, where fortitude's slope in the unarmored data is actually negative!

Recall the result of the linear, unarmored model:

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(linear.fit.level.noarmor)
```

Can this be true? Other influence models, such as GBMs, also show the same thing. Still, it is not an intuitive finding. However, note that:

1. It would explain the existence of high-fortitude, unarmored CL 10s with lots of vulnerabilities -- maximizing fortitude to be as close to 499 as possible would _minimize_ its contribution to creature level!
2. It would explain the question of how many levels armor adds -- some players said 10, others said 6. In this view, obtaining armor does not strictly add a constant creature level; it rolls you into a different segment of the linear model, where the influence of fortitude becomes sharply positive
3. It would also explain why change point detection was so adamant about a change point existing at $fortitude = 500$ -- it's not that armor adds a constant level, it's that parameters of the level calculation change entirely!

I'm unsure if this could be accurate, but it is compelling nonetheless.

## Logarithmic Model Analysis

Further investigation reveals that a **logarithmic transformation** of level provides better fit, suggesting the creature level formula may be multiplicative rather than additive.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Feature engineering for log models
normalized_df <- normalized_df %>%
  mutate(
    kinen = (kinetic + energy) / 2,
    nonkinen = (blast + heat + cold + electricity + acid + stun) / 6
  )

no_armor_df <- normalized_df %>% filter(armor == 0)
armor_df <- normalized_df %>% filter(armor == 1 & fortitude >= 500)
```

Testing `log(level)` as the outcome with raw attributes:

```{r echo=TRUE, message=FALSE, warning=FALSE}
model.log.unarmored <- lm(
  log(level) ~ hardiness + fortitude + dexterity + intellect +
               cleverness + power + courage +
               kinen + nonkinen,
  data = no_armor_df
)

model.log.armored <- lm(
  log(level) ~ hardiness + fortitude + dexterity + intellect +
               cleverness + power + courage +
               kinen + nonkinen,
  data = armor_df
)
```

### Model Fit Comparison

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Calculate R² on original scale
calc_r2_exp <- function(model, data) {
  pred <- exp(predict(model, newdata = data))
  actual <- data$level
  1 - sum((actual - pred)^2) / sum((actual - mean(actual))^2)
}

cat("Log(level) model R² (original scale):\n")
cat("  Unarmored:", round(calc_r2_exp(model.log.unarmored, no_armor_df), 4), "\n")
cat("  Armored:  ", round(calc_r2_exp(model.log.armored, armor_df), 4), "\n")
```

The armored model achieves R² ≈ 0.98, while unarmored achieves R² ≈ 0.94.

### Coefficient Comparison: The Two Formulas

The most striking finding is that **armored and unarmored creatures use fundamentally different formulas**. Several coefficients flip sign between the two regimes:

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Extract coefficients
coef_unarmored <- coef(model.log.unarmored)
coef_armored <- coef(model.log.armored)

# Build comparison table
terms <- c("hardiness", "fortitude", "dexterity", "intellect",
           "cleverness", "power", "courage", "kinen", "nonkinen")

comparison_df <- data.frame(
  Attribute = terms,
  Unarmored = sapply(terms, function(t) {
    val <- coef_unarmored[t]
    sprintf("%+.1f%%", (exp(val * 100) - 1) * 100)
  }),
  Armored = sapply(terms, function(t) {
    val <- coef_armored[t]
    sprintf("%+.1f%%", (exp(val * 100) - 1) * 100)
  }),
  SignFlip = sapply(terms, function(t) {
    ifelse(sign(coef_unarmored[t]) != sign(coef_armored[t]), "YES", "")
  })
)
rownames(comparison_df) <- NULL

knitr::kable(comparison_df,
             col.names = c("Attribute", "Unarmored (per 100)", "Armored (per 100)", "Sign Flip?"),
             caption = "Percentage effect on level per 100 attribute points")
```

### Key Findings

**Sign Flips:**

1. **Fortitude**: The most dramatic flip. For unarmored creatures, fortitude has a *negative* effect on level (-11.7% per 100), while for armored creatures it has a *positive* effect (+5.5% per 100). This is statistically significant in both models.

2. **Hardiness** and **Courage**: Also flip sign, but the armored coefficients are not statistically significant (p > 0.4), so these may be noise.

**Coefficient Magnitude Differences:**

The `kinen` (kinetic/energy resist) coefficient is **3.4x larger** for unarmored creatures (+94.9%) compared to armored (+28.2%). This makes intuitive sense: unarmored creatures rely on resists for survivability, so high resists dramatically increase their effective power level.

### Visualization

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Add predictions
no_armor_df$pred <- exp(predict(model.log.unarmored))
armor_df$pred <- exp(predict(model.log.armored))

# Combine for plotting
no_armor_df$type <- "Unarmored"
armor_df$type <- "Armored"
combined <- bind_rows(
  no_armor_df %>% select(level, pred, type),
  armor_df %>% select(level, pred, type)
)

ggplot(combined, aes(x = pred, y = level, color = type)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  scale_color_manual(values = c("Armored" = "blue", "Unarmored" = "red")) +
  labs(
    title = "Log(Level) Model: Predicted vs Actual",
    subtitle = "Separate formulas for armored and unarmored creatures",
    x = "Predicted Level",
    y = "Actual Level",
    color = "Type"
  ) +
  theme_minimal()
```

### Interpretation

The evidence strongly supports the hypothesis that **the game uses two completely different formulas** for calculating creature level:

**Unarmored Formula** (more complex):

- Base level ≈ 10.6
- Fortitude *lowers* level (bio-engineers could maximize fortitude to ~499 while keeping level low)
- Kinetic/Energy resists have *huge* impact on level
- Offensive attributes (cleverness, power) have larger coefficients

**Armored Formula** (simpler, tighter fit):

- Base level ≈ 12.8
- Fortitude *raises* level (as intuitively expected)
- Resists have moderate impact
- More balanced coefficients across attributes

This explains several long-standing mysteries:

1. **High-fortitude, low-level unarmored creatures existed** because maximizing fortitude actually minimized its contribution to level
2. **Players disagreed on "how many levels armor adds"** because armor doesn't add a constant—it changes the entire formula
3. **The segmentation analysis found a breakpoint at fortitude=500** because crossing that threshold doesn't just add armor, it switches which formula the game uses

### Model Summaries

**Unarmored Model:**

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(model.log.unarmored)
```

**Armored Model:**

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(model.log.armored)
```

## Deriving Implementable Formulas

The log-level models above are useful for understanding the relative importance of attributes, but for re-implementing the game's creature level calculation, we need **linear formulas** that can be directly coded.

### Linear Model Derivation

We fit linear models predicting `level` (not `log(level)`) from raw attributes:

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Linear models for formula derivation
model.armor.linear <- lm(
  level ~ hardiness + fortitude + dexterity + intellect +
          cleverness + power + kinen + nonkinen,
  data = armor_df
)

model.noarmor.linear <- lm(
  level ~ hardiness + fortitude + dexterity + intellect +
          cleverness + power + kinen + nonkinen,
  data = no_armor_df
)
```

### Regression-Derived Coefficients

These are the "raw" coefficients from linear regression, rounded to 3 decimal places:

**Armored Formula (fortitude >= 500):**

```{r echo=FALSE, message=FALSE, warning=FALSE}
coefs.armor <- coef(model.armor.linear)
cat("level =", round(coefs.armor["(Intercept)"], 0), "\n")
cat("      +", round(coefs.armor["hardiness"], 3), "* hardiness\n")
cat("      +", round(coefs.armor["fortitude"], 3), "* fortitude\n")
cat("      +", round(coefs.armor["dexterity"], 3), "* dexterity\n")
cat("      +", round(coefs.armor["intellect"], 3), "* intellect\n")
cat("      +", round(coefs.armor["cleverness"], 3), "* cleverness\n")
cat("      +", round(coefs.armor["power"], 3), "* power\n")
cat("      +", round(coefs.armor["kinen"], 2), "* kinen\n")
cat("      +", round(coefs.armor["nonkinen"], 2), "* nonkinen\n")
```

**Unarmored Formula (fortitude < 500):**

```{r echo=FALSE, message=FALSE, warning=FALSE}
coefs.noarmor <- coef(model.noarmor.linear)
cat("level =", round(coefs.noarmor["(Intercept)"], 0), "\n")
cat("      +", round(coefs.noarmor["hardiness"], 3), "* hardiness\n")
cat("      +", round(coefs.noarmor["fortitude"], 3), "* fortitude  // NEGATIVE!\n")
cat("      +", round(coefs.noarmor["dexterity"], 3), "* dexterity\n")
cat("      +", round(coefs.noarmor["intellect"], 3), "* intellect\n")
cat("      +", round(coefs.noarmor["cleverness"], 3), "* cleverness\n")
cat("      +", round(coefs.noarmor["power"], 3), "* power\n")
cat("      +", round(coefs.noarmor["kinen"], 2), "* kinen\n")
cat("      +", round(coefs.noarmor["nonkinen"], 2), "* nonkinen\n")
```

### Clean Game-Dev Friendly Coefficients

For actual implementation, we want coefficients that represent simple fractions (like 1/100, 1/50, 3/200) that a game developer might reasonably use. These "clean" coefficients are derived by rounding to values that correspond to nice fractions while minimizing loss of fit.

**Armored Formula (Clean):**

```
level = -23
      + 0.01   * hardiness      // 1/100
      + 0.06   * fortitude      // 3/50
      + 0.005  * dexterity      // 1/200
      + 0.01   * intellect      // 1/100
      + 0.025  * cleverness     // 1/40
      + 0.015  * power          // 3/200
      + 0.1    * kinen          // 1/10
      + 0.08   * nonkinen       // 2/25
```

**Unarmored Formula (Clean):**

```
level = 9
      + 0.01   * hardiness      // 1/100
      - 0.02   * fortitude      // -1/50 (NEGATIVE!)
      + 0.01   * dexterity      // 1/100
      + 0.01   * intellect      // 1/100
      + 0.025  * cleverness     // 1/40
      + 0.015  * power          // 3/200
      + 0.12   * kinen          // 3/25
      + 0.06   * nonkinen       // 3/50
```

### Formula Validation

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Helper function to test a formula
test_formula <- function(df, intercept, hardiness, fortitude, dexterity, intellect,
                         cleverness, power, kinen, nonkinen, name) {
  pred <- intercept +
    hardiness * df$hardiness +
    fortitude * df$fortitude +
    dexterity * df$dexterity +
    intellect * df$intellect +
    cleverness * df$cleverness +
    power * df$power +
    kinen * df$kinen +
    nonkinen * df$nonkinen

  residual <- df$level - pred

  ss_res <- sum(residual^2)
  ss_tot <- sum((df$level - mean(df$level))^2)
  r2 <- 1 - ss_res / ss_tot

  return(list(r2 = r2, sd = sd(residual), mean = mean(residual),
              max = max(abs(residual)), pred = pred, residual = residual))
}

# Test regression-derived formulas
armor_reg <- test_formula(armor_df,
  intercept = round(coefs.armor["(Intercept)"], 0),
  hardiness = round(coefs.armor["hardiness"], 3),
  fortitude = round(coefs.armor["fortitude"], 3),
  dexterity = round(coefs.armor["dexterity"], 3),
  intellect = round(coefs.armor["intellect"], 3),
  cleverness = round(coefs.armor["cleverness"], 3),
  power = round(coefs.armor["power"], 3),
  kinen = round(coefs.armor["kinen"], 2),
  nonkinen = round(coefs.armor["nonkinen"], 2),
  "Armored (regression)")

noarmor_reg <- test_formula(no_armor_df,
  intercept = round(coefs.noarmor["(Intercept)"], 0),
  hardiness = round(coefs.noarmor["hardiness"], 3),
  fortitude = round(coefs.noarmor["fortitude"], 3),
  dexterity = round(coefs.noarmor["dexterity"], 3),
  intellect = round(coefs.noarmor["intellect"], 3),
  cleverness = round(coefs.noarmor["cleverness"], 3),
  power = round(coefs.noarmor["power"], 3),
  kinen = round(coefs.noarmor["kinen"], 2),
  nonkinen = round(coefs.noarmor["nonkinen"], 2),
  "Unarmored (regression)")

# Test clean formulas
armor_clean <- test_formula(armor_df,
  intercept = -23,
  hardiness = 0.01,
  fortitude = 0.06,
  dexterity = 0.005,
  intellect = 0.01,
  cleverness = 0.025,
  power = 0.015,
  kinen = 0.1,
  nonkinen = 0.08,
  "Armored (clean)")

noarmor_clean <- test_formula(no_armor_df,
  intercept = 9,
  hardiness = 0.01,
  fortitude = -0.02,
  dexterity = 0.01,
  intellect = 0.01,
  cleverness = 0.025,
  power = 0.015,
  kinen = 0.12,
  nonkinen = 0.06,
  "Unarmored (clean)")

# Build comparison table
validation_df <- data.frame(
  Formula = c("Armored (regression)", "Armored (clean)",
              "Unarmored (regression)", "Unarmored (clean)"),
  R2 = c(armor_reg$r2, armor_clean$r2, noarmor_reg$r2, noarmor_clean$r2),
  SD = c(armor_reg$sd, armor_clean$sd, noarmor_reg$sd, noarmor_clean$sd),
  Mean_Residual = c(armor_reg$mean, armor_clean$mean, noarmor_reg$mean, noarmor_clean$mean),
  Max_Residual = c(armor_reg$max, armor_clean$max, noarmor_reg$max, noarmor_clean$max)
)

knitr::kable(validation_df,
             col.names = c("Formula", "R²", "SD", "Mean Residual", "Max |Residual|"),
             digits = c(0, 4, 2, 2, 2),
             caption = "Formula Validation Metrics")
```

The clean formulas maintain excellent fit:

- **Armored**: R² ≈ 0.98, residuals typically within ±2 levels
- **Unarmored**: R² ≈ 0.91, residuals typically within ±3 levels

### Coefficient Comparison Table

```{r echo=FALSE, message=FALSE, warning=FALSE}
comparison_df <- data.frame(
  Term = c("intercept", "hardiness", "fortitude", "dexterity", "intellect",
           "cleverness", "power", "kinen", "nonkinen"),
  Armor_Reg = c(round(coefs.armor["(Intercept)"], 0),
                round(coefs.armor["hardiness"], 3),
                round(coefs.armor["fortitude"], 3),
                round(coefs.armor["dexterity"], 3),
                round(coefs.armor["intellect"], 3),
                round(coefs.armor["cleverness"], 3),
                round(coefs.armor["power"], 3),
                round(coefs.armor["kinen"], 2),
                round(coefs.armor["nonkinen"], 2)),
  Armor_Clean = c(-23, 0.01, 0.06, 0.005, 0.01, 0.025, 0.015, 0.1, 0.08),
  Unarmor_Reg = c(round(coefs.noarmor["(Intercept)"], 0),
                  round(coefs.noarmor["hardiness"], 3),
                  round(coefs.noarmor["fortitude"], 3),
                  round(coefs.noarmor["dexterity"], 3),
                  round(coefs.noarmor["intellect"], 3),
                  round(coefs.noarmor["cleverness"], 3),
                  round(coefs.noarmor["power"], 3),
                  round(coefs.noarmor["kinen"], 2),
                  round(coefs.noarmor["nonkinen"], 2)),
  Unarmor_Clean = c(9, 0.01, -0.02, 0.01, 0.01, 0.025, 0.015, 0.12, 0.06)
)

knitr::kable(comparison_df,
             col.names = c("Term", "Armored (reg)", "Armored (clean)",
                          "Unarmored (reg)", "Unarmored (clean)"),
             caption = "Coefficient Comparison: Regression vs Clean Values")
```

### Visualization of Clean Formulas

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
# Add clean predictions
armor_df$pred_clean <- armor_clean$pred
no_armor_df$pred_clean <- noarmor_clean$pred

armor_df$type <- "Armored"
no_armor_df$type <- "Unarmored"

combined_clean <- bind_rows(
  armor_df %>% select(level, pred_clean, type),
  no_armor_df %>% select(level, pred_clean, type)
)

ggplot(combined_clean, aes(x = pred_clean, y = level, color = type)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  scale_color_manual(values = c("Armored" = "blue", "Unarmored" = "red")) +
  labs(
    title = "Clean Formulas: Predicted vs Actual Level",
    subtitle = "Using game-dev friendly coefficients",
    x = "Predicted Level",
    y = "Actual Level",
    color = "Type"
  ) +
  theme_minimal()
```

## Final Implementation

The following pseudocode can be used to implement creature level calculation:

```c
int calculateCreatureLevel(Creature* c) {
    // Compute resist averages
    float kinen = (c->kinetic + c->energy) / 2.0f;
    float nonkinen = (c->blast + c->heat + c->cold +
                      c->electricity + c->acid + c->stun) / 6.0f;

    float level;

    if (c->fortitude >= 500) {
        // ARMORED FORMULA
        level = -23.0f
              + 0.01f   * c->hardiness
              + 0.06f   * c->fortitude
              + 0.005f  * c->dexterity
              + 0.01f   * c->intellect
              + 0.025f  * c->cleverness
              + 0.015f  * c->power
              + 0.1f    * kinen
              + 0.08f   * nonkinen;
    } else {
        // UNARMORED FORMULA (note: fortitude is SUBTRACTED)
        level = 9.0f
              + 0.01f   * c->hardiness
              - 0.02f   * c->fortitude    // NEGATIVE!
              + 0.01f   * c->dexterity
              + 0.01f   * c->intellect
              + 0.025f  * c->cleverness
              + 0.015f  * c->power
              + 0.12f   * kinen
              + 0.06f   * nonkinen;
    }

    // Clamp and round
    if (level < 1.0f) level = 1.0f;
    if (level > 75.0f) level = 75.0f;
    return (int)(level + 0.5f);  // round to nearest
}
```

### Key Implementation Notes

1. **Armor threshold**: The switch point is `fortitude >= 500`, which corresponds to the point where creatures gain armor
2. **Fortitude sign flip**: The most dramatic finding—fortitude has opposite effects in the two formulas
3. **kinen vs nonkinen**: Kinetic/Energy resists (capped at 60%) are weighted differently than other resists
4. **Residual sources**: Remaining error is likely due to rounding in the original game's attribute calculations

