---
title: "Creature Level Analysis"
output: html_document
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("R/data.R")
library(mgcv)
library(broom)
library(ggplot2)
library(lmtest)
library(segmented)

no_armor_df <- normalized_df %>% dplyr::filter(armor == 0)
armor_df    <- normalized_df %>% dplyr::filter(armor == 1 & fortitude >= 500)
```

# Overview

Bio-Engineering in Star Wars: Galaxies let players craft custom creatures
("pets") by combining DNA samples. The creature's level (CL) governed who
could tame the pet and how powerful it was. The 2003 BE Guide noted that CL
"mostly depends on damage, resists, and the CL of the donors," but the exact
algorithm was never published.

This document reverse-engineers that algorithm from the Furrycat archive.
The headline finding: **the game uses two completely different formulas**,
selected by whether the creature has armor (`fortitude >= 500`). Both
formulas are linear in the creature's attributes, but one of them — the
unarmored formula — needs a vulnerability-aware resist term to fit the data
well.

The dataset, after filtering, has 370 creatures: 79 armored, 291 unarmored.

# An Initial GAM

A single GAM across all creatures gives a quick picture of which attributes
move level.

```{r echo=TRUE, message=FALSE, warning=FALSE}
model.gam.everything <- gam(
  level ~
    s(hardiness) +
    s(fortitude) +
    s(dexterity) +
    s(endurance) +
    s(intellect) +
    s(cleverness) +
    s(dependability) +
    s(courage) +
    s(fierceness) +
    s(power) +
    s(kinetic) +
    s(energy)  +
    s(blast) +
    s(heat) +
    s(cold) +
    s(electricity) +
    s(acid) +
    s(stun),
  data = normalized_df,
  family = gaussian()
)
summary(model.gam.everything)
```

Several attributes appear important:

- hardiness, fortitude, dexterity, intellect, cleverness, power
- kinetic, energy (and to a lesser extent cold)

And a handful land on ~1.00 effective degrees of freedom (essentially flat):
endurance, fierceness, blast, heat, electricity. We treat those as noise
attributes for level purposes.

The fortitude smooth is the most interesting term. Its shape will turn out
to encode the two-formula structure.

# Feature Engineering

To avoid over-fitting to small per-attribute weight differences, we collapse
correlated attributes into synthetic features:

1. `average_hdi = (hardiness + dexterity + intellect) / 3` — these three
   attributes have similar weights against level; collapsing them prevents
   the model from over-attributing influence to whichever happens to
   correlate most with fortitude.
2. `kinen = (kinetic + energy) / 2` — kinetic and energy share the in-game
   60% cap; treating them as one resist class respects that mechanic.
3. `nonkinen = (blast + heat + cold + electricity + acid + stun) / 6` —
   the other six resists, none of which are individually significant after
   collapsing.

These are derived in `R/data.R` and ride along on `normalized_df`.

A re-fit GAM with these collapsed features reproduces the same picture
with far fewer terms:

```{r echo=TRUE, message=FALSE, warning=FALSE}
model.gam <- gam(
  level ~
    s(average_hdi) +
    s(fortitude) +
    s(cleverness) +
    s(power) +
    s(kinen) +
    s(nonkinen),
  data = normalized_df
)
summary(model.gam)
```

Holding all other attributes at their mean, the fortitude smooth shows the
discontinuity that motivates the two-formula split:

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4}
plot(model.gam, select = 2, shade = TRUE,
     main = "Fortitude smooth (level vs fortitude, all else fixed)")
abline(v = 500, col = "red", lty = 2)
```

The smooth is **decreasing** below fortitude=500 and **increasing** above
it. We'll quantify and verify that breakpoint shortly.

# Armored Creatures

For creatures with `fortitude >= 500`, a plain linear regression already
fits remarkably well.

```{r echo=TRUE, message=FALSE, warning=FALSE}
linear.fit.level.armor <- lm(
  level ~ average_hdi + fortitude + cleverness + power + kinen + nonkinen,
  data = armor_df
)
summary(linear.fit.level.armor)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(armor_df,
       aes(x = predict(linear.fit.level.armor, newdata = armor_df),
           y = level)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Armored: predicted vs actual level",
       x = "predicted_level", y = "level") +
  theme_minimal()
```

Residual diagnostics:

```{r echo=FALSE, message=FALSE, warning=FALSE}
rs.armor <- resid(linear.fit.level.armor)
qqnorm(rs.armor); qqline(rs.armor)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
shapiro.test(rs.armor)
bptest(linear.fit.level.armor)
```

The armored residuals are small (±4 levels), normally distributed, and
homoscedastic. This is consistent with the residuals being driven by
crafting-system randomness rather than a missing variable.

A clean, game-dev-friendly version of the armored formula:

```
level = -23
      + 0.01   * hardiness
      + 0.06   * fortitude
      + 0.005  * dexterity
      + 0.01   * intellect
      + 0.025  * cleverness
      + 0.015  * power
      + 0.10   * kinen
      + 0.08   * nonkinen
```

R² ≈ 0.98, residual SD ≈ 1.8 levels.

# Unarmored Creatures: the Naive Fit

The same recipe on unarmored creatures gives a worse fit, and the residuals
clearly violate normality.

```{r echo=TRUE, message=FALSE, warning=FALSE}
linear.fit.level.noarmor <- lm(
  level ~ average_hdi + fortitude + cleverness + power + kinen + nonkinen,
  data = no_armor_df
)
summary(linear.fit.level.noarmor)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(no_armor_df,
       aes(x = predict(linear.fit.level.noarmor, newdata = no_armor_df),
           y = level)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Unarmored (naive `kinen`): predicted vs actual",
       x = "predicted_level", y = "level") +
  theme_minimal()
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
rs.noarmor <- resid(linear.fit.level.noarmor)
qqnorm(rs.noarmor); qqline(rs.noarmor)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
shapiro.test(rs.noarmor)
bptest(linear.fit.level.noarmor)
```

Three things stand out from this naive unarmored fit:

1. **Fortitude has a negative coefficient** in unarmored creatures. This is
   reproducible across modelling approaches. We come back to it below.
2. The **residuals are not normal**. The Q-Q plot shows a heavy left tail —
   creatures that the model over-predicts.
3. The over-predicted creatures cluster at low CL, with a recognizable
   pattern that isn't captured by the naive `kinen` average.

# The "Uber CL10" Cluster

A heavy left tail in the residuals turned out to be a coherent cluster:
low-CL unarmored creatures with ~10k Health and a kinetic resist near the
60% cap.

```{r echo=FALSE, message=FALSE, warning=FALSE}
no_armor_df <- no_armor_df %>%
  mutate(naive_resid = level - predict(linear.fit.level.noarmor, newdata = .))

uber_cl10 <- no_armor_df %>%
  dplyr::filter(level <= 12, kinetic >= 50)

cat("Uber-CL10-style cluster (level <= 12 and kinetic >= 50):\n")
cat(sprintf("  n = %d\n", nrow(uber_cl10)))
cat(sprintf("  mean kinetic = %.1f, mean energy = %.1f\n",
            mean(uber_cl10$kinetic), mean(uber_cl10$energy)))
cat(sprintf("  mean kinen = (k+e)/2 = %.1f\n", mean(uber_cl10$kinen)))
cat(sprintf("  mean naive residual = %.2f (negative = over-predicted)\n",
            mean(uber_cl10$naive_resid)))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5}
no_armor_df <- no_armor_df %>%
  mutate(uber = ifelse(level <= 12 & kinetic >= 50, "uber CL10", "other"))

ggplot(no_armor_df, aes(x = kinetic, y = energy, color = uber)) +
  geom_point(alpha = 0.7) +
  geom_hline(yintercept = 0, color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 60, color = "grey50", linetype = "dashed") +
  scale_color_manual(values = c("uber CL10" = "red", "other" = "grey60")) +
  labs(title = "Unarmored creatures: kinetic vs energy resist",
       subtitle = "Uber-CL10 cluster has kinetic near 60% cap, energy heavily negative",
       x = "kinetic", y = "energy") +
  theme_minimal()
```

The Bio-Engineer "trick": stack kinetic resist near the 60% cap, accept a
heavy negative energy resist as the cost. The naive `kinen = (k+e)/2`
averages a near-cap positive against a heavily-negative number and credits
the small remainder linearly — so the formula thinks the creature has ~5%
"average" resist and predicts CL ~13, when in practice the energy
vulnerability erases the resist's value entirely. The game evidently
penalizes the vulnerability harder than `(k+e)/2` does.

# Vulnerability-Aware Resists

We test three symmetric formulations of the kinetic/energy term against
the `kinen` baseline. All four use the same other predictors; only the
resist term changes.

| ID | Resist term | Mechanic |
|---|---|---|
| M0 | `kinen = (k+e)/2` | naive average |
| M5 | `kinen_pos + kinen_neg` (split positive / negative half) | asymmetric credit on positive vs penalty on negative |
| M6 | `pmax(kinen, 0)` | average resist, but no negative credit |
| M7 | `pmax(pmin(k, e), 0)` | use the **weaker** side, floored at zero |

Symmetric in kinetic and energy — both share the 60% in-engine cap, so
treating them as a single resist class is mechanically motivated.

```{r echo=TRUE, message=FALSE, warning=FALSE}
no_armor_df <- no_armor_df %>%
  mutate(
    kinen_pos = pmax(kinen, 0),
    kinen_neg = pmin(kinen, 0),
    ke_floor  = pmax(pmin(kinetic, energy), 0)
  )

m0 <- lm(level ~ hardiness + fortitude + dexterity + intellect +
                 cleverness + power + kinen + nonkinen,
         data = no_armor_df)
m5 <- lm(level ~ hardiness + fortitude + dexterity + intellect +
                 cleverness + power + kinen_pos + kinen_neg + nonkinen,
         data = no_armor_df)
m6 <- lm(level ~ hardiness + fortitude + dexterity + intellect +
                 cleverness + power + kinen_pos + nonkinen,
         data = no_armor_df)
m7 <- lm(level ~ hardiness + fortitude + dexterity + intellect +
                 cleverness + power + ke_floor + nonkinen,
         data = no_armor_df)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
uber_idx <- no_armor_df$level <= 12 & no_armor_df$kinetic >= 50

cluster_sd <- function(m) {
  sd(no_armor_df$level[uber_idx] - predict(m)[uber_idx])
}

resist_compare <- data.frame(
  Model = c("M0 kinen", "M5 kinen_pos+kinen_neg", "M6 pmax(kinen,0)",
            "M7 pmax(pmin(k,e),0)"),
  R2    = c(summary(m0)$r.squared, summary(m5)$r.squared,
            summary(m6)$r.squared, summary(m7)$r.squared),
  ResidSD = c(sd(resid(m0)), sd(resid(m5)), sd(resid(m6)), sd(resid(m7))),
  AIC   = c(AIC(m0), AIC(m5), AIC(m6), AIC(m7)),
  UberCL10_SD = c(cluster_sd(m0), cluster_sd(m5), cluster_sd(m6), cluster_sd(m7))
)
knitr::kable(resist_compare, digits = c(0, 4, 3, 1, 3),
             col.names = c("Model", "R²", "Resid SD", "AIC", "Uber-CL10 SD"),
             caption = "Resist-term variants on the unarmored data")
```

M7 is unambiguously best on every metric: highest R², lowest residual SD,
lowest AIC, and crucially the tightest fit on the uber-CL10 cluster
itself (SD 0.70 vs 1.10 baseline). M5 and M6 capture the cluster's mean
but not its tightness — a creature with `kinetic = 60`, `energy = -50`
still gets `pmax(kinen, 0) = 5` of credit it apparently shouldn't.

The mechanism reading: **vulnerability on either kinetic or energy erases
all kin/eng resist credit**. Roughly half of unarmored creatures get
zero credit from this term because they have a vulnerability on one
side.

We promote M7 as the canonical unarmored form. The fitted coefficients:

```{r echo=TRUE, message=FALSE, warning=FALSE}
summary(m7)
```

Q-Q residuals:

```{r echo=FALSE, message=FALSE, warning=FALSE}
qqnorm(resid(m7)); qqline(resid(m7))
```

The Shapiro-Wilk test still rejects normality (p ≈ 3e-6), so there is
additional structure left in the unarmored residuals beyond the
vulnerability mechanism — see "Persistent Outliers" below.

# Verifying the fortitude=500 Breakpoint

The two-formula split assumes a hard discontinuity at fortitude=500. We
test it four ways: a single global linear regression with no break, a
GAM smooth, a data-driven `segmented` regression, and the proposed
hard split.

```{r echo=TRUE, message=FALSE, warning=FALSE}
df <- normalized_df %>%
  mutate(ke_floor = pmax(pmin(kinetic, energy), 0))

# Single global linear (no break)
m_global <- lm(level ~ hardiness + fortitude + dexterity + intellect +
                       cleverness + power + ke_floor + nonkinen,
               data = df)

# GAM with smooth fortitude
m_gam <- gam(level ~ s(fortitude) + hardiness + dexterity + intellect +
                     cleverness + power + ke_floor + nonkinen,
             data = df)

# Segmented regression (data-driven breakpoint)
seg_in <- lm(level ~ fortitude + hardiness + dexterity + intellect +
                     cleverness + power + ke_floor + nonkinen, data = df)
m_seg  <- segmented(seg_in, seg.Z = ~ fortitude, psi = 500)

# Two formulas, hard split at fort = 500
arm_df <- df %>% dplyr::filter(armor == 1 & fortitude >= 500)
una_df <- df %>% dplyr::filter(armor == 0)
m_arm  <- lm(level ~ hardiness + fortitude + dexterity + intellect +
                     cleverness + power + ke_floor + nonkinen,
             data = arm_df)
m_una  <- lm(level ~ hardiness + fortitude + dexterity + intellect +
                     cleverness + power + ke_floor + nonkinen,
             data = una_df)

resid_split <- c(resid(m_arm), resid(m_una))
ss_split    <- sum(resid_split^2)
n_split     <- length(resid_split)
r2_split    <- 1 - ss_split / sum((c(arm_df$level, una_df$level) -
                                   mean(c(arm_df$level, una_df$level)))^2)
sd_split    <- sd(resid_split)
# AIC for the joint two-formula model: sum of AICs (each is independent)
aic_split   <- AIC(m_arm) + AIC(m_una)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
breakpoint_compare <- data.frame(
  Model = c("Single global linear (no break)",
            "GAM smooth fortitude",
            sprintf("Segmented regression (break at fort=%.0f)",
                    m_seg$psi[1, "Est."]),
            "Two formulas, hard split at fort=500"),
  R2    = c(summary(m_global)$r.squared,
            summary(m_gam)$r.sq,
            summary(m_seg)$r.squared,
            r2_split),
  ResidSD = c(sd(resid(m_global)), sd(resid(m_gam)),
              sd(resid(m_seg)), sd_split),
  AIC   = c(AIC(m_global), AIC(m_gam), AIC(m_seg), aic_split)
)
knitr::kable(breakpoint_compare, digits = c(0, 4, 3, 1),
             col.names = c("Model", "R²", "Resid SD", "AIC"),
             caption = "Breakpoint test (level ~ stats with different fortitude treatments)")
```

The hard split at fortitude=500 beats the single global linear by ~500
AIC points and beats a flexible GAM smooth by ~50. That second margin is
the important one: it says the relationship has a true *discontinuity*
at 500, not just a slope change. Classic step function in the game code.

The sign flip across the break is iron-clad:

```{r echo=FALSE, message=FALSE, warning=FALSE}
fort_compare <- data.frame(
  Subset = c("Unarmored (fortitude < 500)", "Armored (fortitude >= 500)"),
  N      = c(nobs(m_una), nobs(m_arm)),
  Fortitude_coef = c(coef(m_una)["fortitude"], coef(m_arm)["fortitude"]),
  P_value = c(summary(m_una)$coef["fortitude", "Pr(>|t|)"],
              summary(m_arm)$coef["fortitude", "Pr(>|t|)"])
)
knitr::kable(fort_compare, digits = c(0, 0, 5, -1),
             col.names = c("Subset", "n", "Fortitude coef", "p-value"),
             caption = "Fortitude coefficient on either side of the break")
```

Both coefficients are extremely significant, and they have opposite signs.

# The Negative-Fortitude Coefficient: Mechanism vs Artifact

The unarmored fortitude coefficient is **negative**. Taken at face value,
this would say that a creature with more fortitude has a *lower* level —
counterintuitive enough to deserve scrutiny. We tested two
explanations: a stat-budget / collinearity artifact, and an
effective-resist double-count.

## Stat-budget / collinearity diagnosis

Within the unarmored subset, fortitude correlates 0.85 with hardiness and
0.84 with health — Bio-Engineers wire these up together when crafting.
Drop fortitude from the model, and the hardiness coefficient changes
dramatically:

```{r echo=TRUE, message=FALSE, warning=FALSE}
m_una_full <- m_una
m_una_no_fort <- lm(level ~ hardiness + dexterity + intellect +
                            cleverness + power + ke_floor + nonkinen,
                    data = una_df)

shift_compare <- data.frame(
  Term = c("hardiness", "dexterity", "intellect",
           "cleverness", "power", "ke_floor", "nonkinen"),
  With_fortitude = coef(m_una_full)[c("hardiness","dexterity","intellect",
                                      "cleverness","power","ke_floor","nonkinen")],
  Without_fortitude = coef(m_una_no_fort)[c("hardiness","dexterity","intellect",
                                            "cleverness","power","ke_floor","nonkinen")]
)
shift_compare$Shift <- shift_compare$Without_fortitude - shift_compare$With_fortitude
knitr::kable(shift_compare, digits = 4, row.names = FALSE,
             caption = "Coefficient shifts when fortitude is dropped from the unarmored model")
```

The hardiness coefficient flips from +0.013 to roughly zero. That is the
smoking gun: the regression had been crediting hardiness positively and
fortitude negatively as a *paired* signal — what it is actually
identifying is the contrast between hardiness and fortitude, not their
independent contributions.

A manual VIF (variance inflation factor) puts numbers on the
collinearity:

```{r echo=TRUE, message=FALSE, warning=FALSE}
manual_vif <- function(model) {
  X <- model.matrix(model)[, -1, drop = FALSE]
  vifs <- sapply(seq_len(ncol(X)), function(j) {
    fit <- lm(X[, j] ~ X[, -j])
    1 / (1 - summary(fit)$r.squared)
  })
  setNames(vifs, colnames(X))
}
round(manual_vif(m_una_full), 2)
```

Fortitude itself is at moderate VIF (~5); the worse offenders are
intellect and cleverness, which collinearity-correlate with the rest of
the stat budget too.

**Mechanistic reading**: the game most plausibly gives unarmored
fortitude *zero* level contribution — without armor, fortitude doesn't
matter for damage mitigation — but BE crafters co-vary fortitude
alongside hardiness and health when optimizing. The regression
discovers an empirical (hardiness +, fortitude −) pair that fits the
data well even though only one of the two is mechanistically real.

**Practical upshot**: keep the empirical coefficients (the formula
predicts well), but don't read literal mechanism into the negative
fortitude term. The "uber CL10" phenomenon's apparent
"fortitude penalty" effect is more plausibly explained by the M7
vulnerability mechanism plus this stat-budget correlation.

## Effective vs special resists: ruled out

A second hypothesis: in the game, kinetic and energy resists come from
two sources — **effective** resists derived directly from fortitude
(`floor(fortitude / 10)` for unarmored) and **special** resists from
DNA samples. When both are present, special wins and effective drops
out. For creatures whose displayed kinetic/energy *is* effective, the
column M7 consumes is literally a function of fortitude — a potential
double-count that could create the spurious coefficient.

We confirmed the formula and tested whether splitting by source
explains anything:

```{r echo=TRUE, message=FALSE, warning=FALSE}
unarm <- normalized_df %>% dplyr::filter(armor == 0) %>%
  mutate(
    ke_floor    = pmax(pmin(kinetic, energy), 0),
    k_effective = !is.na(kinetic.effective) & kinetic.effective != 0,
    e_effective = !is.na(energy.effective)  & energy.effective  != 0,
    k_special   = !is.na(kinetic.special)   & kinetic.special   != 0,
    e_special   = !is.na(energy.special)    & energy.special    != 0
  )

# Verify: kinetic.effective = floor(fortitude / 10) for unarmored
unarm_eff <- unarm %>% dplyr::filter(k_effective)
exact_match <- sum(unarm_eff$kinetic.effective == floor(unarm_eff$fortitude / 10))
cat(sprintf("kinetic.effective matches floor(fortitude/10): %d / %d\n",
            exact_match, nrow(unarm_eff)))

# Refit on the all-special subset (no fortitude->resist algebraic relationship)
all_special <- unarm %>%
  dplyr::filter(!k_effective & !e_effective &
                (k_special | kinetic == 0) &
                (e_special | energy == 0))
m_special <- lm(level ~ hardiness + fortitude + dexterity + intellect +
                        cleverness + power + ke_floor + nonkinen,
                data = all_special)

cat(sprintf("Full unarmored fortitude coef:  %.5f (n=%d)\n",
            coef(m_una)["fortitude"], nobs(m_una)))
cat(sprintf("All-special  fortitude coef:    %.5f (n=%d)\n",
            coef(m_special)["fortitude"], nobs(m_special)))
```

The fortitude coefficient is essentially identical on the all-special
subset, where there is *no* algebraic relationship between fortitude
and the kinetic/energy values M7 consumes. The effective-resist
double-count is ruled out as the cause of the negative coefficient.
The stat-budget diagnosis stands.

(For completeness: zeroing out effective resists from `ke_floor` —
treating the game as if it ignored them — makes the fit worse by ~40
AIC. The level formula appears to credit effective and special resists
at the same rate, using whichever value is displayed. So M7 is using
the right column; no source-aware refactor is needed.)

# Combined Fit: Both Formulas

Put the two formulas together with a hard switch at fortitude=500:

```{r echo=TRUE, message=FALSE, warning=FALSE}
source("R/creature_level_model.R")

normalized_df$predicted_level <- custom_model(normalized_df)
normalized_df$residuals       <- normalized_df$level - normalized_df$predicted_level

cat(sprintf("Combined: n=%d, R^2=%.4f, residual SD=%.3f\n",
            nrow(normalized_df),
            1 - sum(normalized_df$residuals^2) /
                sum((normalized_df$level - mean(normalized_df$level))^2),
            sd(normalized_df$residuals)))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
normalized_df <- normalized_df %>%
  mutate(armor_label = ifelse(armor == 1 & fortitude >= 500,
                              "Armored (fort >= 500)", "Unarmored"))

ggplot(normalized_df, aes(x = predicted_level, y = level, color = armor_label)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, color = "black", linetype = "dashed") +
  scale_color_manual(values = c("Armored (fort >= 500)" = "blue",
                                "Unarmored" = "red")) +
  labs(title = "Two-formula model: predicted vs actual level",
       subtitle = "M7 weaker-resist-floor for unarmored; clean linear for armored",
       x = "Predicted level", y = "Actual level", color = NULL) +
  theme_minimal()
```

# Final Formulas

**Armored (fortitude >= 500)** — clean fractions, R² ≈ 0.98:

```
level = -23
      + 0.01   * hardiness
      + 0.06   * fortitude
      + 0.005  * dexterity
      + 0.01   * intellect
      + 0.025  * cleverness
      + 0.015  * power
      + 0.10   * kinen
      + 0.08   * nonkinen
```

with `kinen = (kinetic + energy) / 2` and
`nonkinen = (blast + heat + cold + electricity + acid + stun) / 6`.

**Unarmored (fortitude < 500)** — M7 weaker-resist-floor, R² ≈ 0.95:

```
level = 7.87
      + 0.013  * hardiness
      - 0.019  * fortitude
      + 0.004  * dexterity
      + 0.010  * intellect
      + 0.025  * cleverness
      + 0.015  * power
      + 0.167  * pmax(pmin(kinetic, energy), 0)
      + 0.054  * nonkinen
```

The kinetic/energy resist term is the **weaker side, floored at zero**
— vulnerability on either kinetic or energy erases all credit for the
kin/eng resist class. Symmetric in kinetic and energy.

Combined (both formulas, n=370): R² ≈ 0.98, residual SD ≈ 1.9 levels.

# Persistent Outliers

After M7, residual non-normality is mild but still present (Shapiro-Wilk
p ≈ 3e-6 on the unarmored set). Two structural patterns survive:

1. **High-DPS under-prediction**: a handful of high-cleverness/power
   creatures (e.g. rancor CL48, falumpaset CL32, razor cat CL49)
   under-predict by 5–7 levels. Earlier scratchpad work suggested
   diminishing-returns thresholds at cleverness 200/300/400 — worth
   revisiting under M7.
2. **Near-armor-border zone**: residuals concentrate at
   fortitude 380–540, hinting that the hard `fortitude < 500` step in
   the implementation may not match the game exactly — possibly a
   transition band the two formulas don't bridge cleanly.

A 10-creature "data-quality suspects" sensitivity analysis (low-
health-for-skin entries plus the 6 effective-resist formula mismatches)
shifts coefficients by less than 0.01 each and improves combined SD
from 1.92 to 1.88. Coefficient stability under that ablation gives us
confidence the formula is not being driven by a small number of
suspect rows.

# Implementation Pseudocode

```c
int calculateCreatureLevel(const Creature* c) {
    float kinen = (c->kinetic + c->energy) / 2.0f;
    float nonkinen = (c->blast + c->heat + c->cold +
                      c->electricity + c->acid + c->stun) / 6.0f;

    float level;

    if (c->fortitude >= 500) {
        // ARMORED FORMULA
        level = -23.0f
              + 0.01f   * c->hardiness
              + 0.06f   * c->fortitude
              + 0.005f  * c->dexterity
              + 0.01f   * c->intellect
              + 0.025f  * c->cleverness
              + 0.015f  * c->power
              + 0.1f    * kinen
              + 0.08f   * nonkinen;
    } else {
        // UNARMORED FORMULA — M7 weaker-resist-floor.
        // ke_floor = max(0, min(kinetic, energy)): if either side has a
        // vulnerability, the kin/eng resist credit collapses to zero.
        float ke_floor = fmaxf(0.0f, fminf(c->kinetic, c->energy));

        level = 7.87f
              + 0.013f  * c->hardiness
              - 0.019f  * c->fortitude
              + 0.004f  * c->dexterity
              + 0.010f  * c->intellect
              + 0.025f  * c->cleverness
              + 0.015f  * c->power
              + 0.167f  * ke_floor
              + 0.054f  * nonkinen;
    }

    if (level < 1.0f)  level = 1.0f;

    return (int)(level + 0.5f);   // round to nearest
}
```

The R counterpart of this pseudocode lives in `R/creature_level_model.R`
as `custom_model()`.
