---
title: "Creature Level Analysis"
output: html_document
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("R/data.R")
library(mgcv)
library(broom)
library(ggplot2)
library(lmtest)
library(segmented)

no_armor_df <- normalized_df %>% dplyr::filter(armor == 0)
armor_df    <- normalized_df %>% dplyr::filter(armor == 1 & fortitude >= 500)
```

# Overview

Bio-Engineering in Star Wars: Galaxies let players craft custom creatures
("pets") by combining DNA samples. The creature's level (CL) governed who
could tame the pet and how powerful it was. The 2003 BE Guide noted that CL
"mostly depends on damage, resists, and the CL of the donors," but the exact
algorithm was never published.

This document reverse-engineers that algorithm from the Furrycat archive.
The headline finding: **the game uses two completely different formulas**,
selected by whether the creature has armor (`fortitude >= 500`). Both
formulas are linear in the creature's attributes, but one of them — the
unarmored formula — needs a vulnerability-aware resist term to fit the data
well.

The dataset, after filtering, has 370 creatures: 79 armored, 291 unarmored.

# An Initial GAM

A single GAM across all creatures gives a quick picture of which attributes
move level.

```{r echo=TRUE, message=FALSE, warning=FALSE}
model.gam.everything <- gam(
  level ~
    s(hardiness) +
    s(fortitude) +
    s(dexterity) +
    s(endurance) +
    s(intellect) +
    s(cleverness) +
    s(dependability) +
    s(courage) +
    s(fierceness) +
    s(power) +
    s(kinetic) +
    s(energy)  +
    s(blast) +
    s(heat) +
    s(cold) +
    s(electricity) +
    s(acid) +
    s(stun),
  data = normalized_df,
  family = gaussian()
)
summary(model.gam.everything)
```

Several attributes appear important:

- hardiness, fortitude, dexterity, intellect, cleverness, power
- kinetic, energy (and to a lesser extent cold)

And a handful land on ~1.00 effective degrees of freedom (essentially flat):
endurance, fierceness, blast, heat, electricity. We treat those as noise
attributes for level purposes.

The fortitude smooth is the most interesting term. Its shape will turn out
to encode the two-formula structure.

# Feature Engineering

To avoid over-fitting to small per-attribute weight differences, we collapse
correlated attributes into synthetic features:

1. `average_hdi = (hardiness + dexterity + intellect) / 3` — these three
   attributes have similar weights against level; collapsing them prevents
   the model from over-attributing influence to whichever happens to
   correlate most with fortitude.
2. `kinen = (kinetic + energy) / 2` — kinetic and energy share the in-game
   60% cap; treating them as one resist class respects that mechanic.
3. `nonkinen = (blast + heat + cold + electricity + acid + stun) / 6` —
   the other six resists, none of which are individually significant after
   collapsing.

These are derived in `R/data.R` and ride along on `normalized_df`.

A re-fit GAM with these collapsed features reproduces the same picture
with far fewer terms:

```{r echo=TRUE, message=FALSE, warning=FALSE}
model.gam <- gam(
  level ~
    s(average_hdi) +
    s(fortitude) +
    s(cleverness) +
    s(power) +
    s(kinen) +
    s(nonkinen),
  data = normalized_df
)
summary(model.gam)
```

Holding all other attributes at their mean, the fortitude smooth shows the
discontinuity that motivates the two-formula split:

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4}
plot(model.gam, select = 2, shade = TRUE,
     main = "Fortitude smooth (level vs fortitude, all else fixed)")
abline(v = 500, col = "red", lty = 2)
```

The smooth is **decreasing** below fortitude=500 and **increasing** above
it. We'll quantify and verify that breakpoint shortly.

# Armored Creatures

For creatures with `fortitude >= 500`, a plain linear regression already
fits remarkably well.

```{r echo=TRUE, message=FALSE, warning=FALSE}
linear.fit.level.armor <- lm(
  level ~ average_hdi + fortitude + cleverness + power + kinen + nonkinen,
  data = armor_df
)
summary(linear.fit.level.armor)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(armor_df,
       aes(x = predict(linear.fit.level.armor, newdata = armor_df),
           y = level)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Armored: predicted vs actual level",
       x = "predicted_level", y = "level") +
  theme_minimal()
```

Residual diagnostics:

```{r echo=FALSE, message=FALSE, warning=FALSE}
rs.armor <- resid(linear.fit.level.armor)
qqnorm(rs.armor); qqline(rs.armor)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
shapiro.test(rs.armor)
bptest(linear.fit.level.armor)
```

The armored residuals are small (±4 levels), normally distributed, and
homoscedastic. This is consistent with the residuals being driven by
crafting-system randomness rather than a missing variable.

A clean, game-dev-friendly version of the armored formula:

```
level = -23
      + 0.01   * hardiness
      + 0.06   * fortitude
      + 0.005  * dexterity
      + 0.01   * intellect
      + 0.025  * cleverness
      + 0.015  * power
      + 0.10   * kinen
      + 0.08   * nonkinen
```

R² ≈ 0.98, residual SD ≈ 1.8 levels.

# Unarmored Creatures: the Naive Fit

The same recipe on unarmored creatures gives a worse fit, and the residuals
clearly violate normality.

```{r echo=TRUE, message=FALSE, warning=FALSE}
linear.fit.level.noarmor <- lm(
  level ~ average_hdi + fortitude + cleverness + power + kinen + nonkinen,
  data = no_armor_df
)
summary(linear.fit.level.noarmor)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(no_armor_df,
       aes(x = predict(linear.fit.level.noarmor, newdata = no_armor_df),
           y = level)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Unarmored (naive `kinen`): predicted vs actual",
       x = "predicted_level", y = "level") +
  theme_minimal()
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
rs.noarmor <- resid(linear.fit.level.noarmor)
qqnorm(rs.noarmor); qqline(rs.noarmor)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
shapiro.test(rs.noarmor)
bptest(linear.fit.level.noarmor)
```

Three things stand out from this naive unarmored fit:

1. **Fortitude has a negative coefficient** in unarmored creatures. This is
   reproducible across modelling approaches. We come back to it below.
2. The **residuals are not normal**. The Q-Q plot shows a heavy left tail —
   creatures that the model over-predicts.
3. The over-predicted creatures cluster at low CL, with a recognizable
   pattern that isn't captured by the naive `kinen` average.

# The "Uber CL10" Cluster

A heavy left tail in the residuals turned out to be a coherent cluster:
low-CL unarmored creatures with ~10k Health and a kinetic resist near the
60% cap.

```{r echo=FALSE, message=FALSE, warning=FALSE}
no_armor_df <- no_armor_df %>%
  mutate(naive_resid = level - predict(linear.fit.level.noarmor, newdata = .))

uber_cl10 <- no_armor_df %>%
  dplyr::filter(level <= 12, kinetic >= 50)

cat("Uber-CL10-style cluster (level <= 12 and kinetic >= 50):\n")
cat(sprintf("  n = %d\n", nrow(uber_cl10)))
cat(sprintf("  mean kinetic = %.1f, mean energy = %.1f\n",
            mean(uber_cl10$kinetic), mean(uber_cl10$energy)))
cat(sprintf("  mean kinen = (k+e)/2 = %.1f\n", mean(uber_cl10$kinen)))
cat(sprintf("  mean naive residual = %.2f (negative = over-predicted)\n",
            mean(uber_cl10$naive_resid)))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5}
no_armor_df <- no_armor_df %>%
  mutate(uber = ifelse(level <= 12 & kinetic >= 50, "uber CL10", "other"))

ggplot(no_armor_df, aes(x = kinetic, y = energy, color = uber)) +
  geom_point(alpha = 0.7) +
  geom_hline(yintercept = 0, color = "grey50", linetype = "dashed") +
  geom_vline(xintercept = 60, color = "grey50", linetype = "dashed") +
  scale_color_manual(values = c("uber CL10" = "red", "other" = "grey60")) +
  labs(title = "Unarmored creatures: kinetic vs energy resist",
       subtitle = "Uber-CL10 cluster has kinetic near 60% cap, energy heavily negative",
       x = "kinetic", y = "energy") +
  theme_minimal()
```

The Bio-Engineer "trick": stack kinetic resist near the 60% cap, accept a
heavy negative energy resist as the cost. The naive `kinen = (k+e)/2`
averages a near-cap positive against a heavily-negative number and credits
the small remainder linearly — so the formula thinks the creature has ~5%
"average" resist and predicts CL ~13, when in practice the energy
vulnerability erases the resist's value entirely. The game evidently
penalizes the vulnerability harder than `(k+e)/2` does.

# Vulnerability-Aware Resists

We test three symmetric formulations of the kinetic/energy term against
the `kinen` baseline. All four use the same other predictors; only the
resist term changes.

| ID | Resist term | Mechanic |
|---|---|---|
| M0 | `kinen = (k+e)/2` | naive average |
| M5 | `kinen_pos + kinen_neg` (split positive / negative half) | asymmetric credit on positive vs penalty on negative |
| M6 | `pmax(kinen, 0)` | average resist, but no negative credit |
| M7 | `pmax(pmin(k, e), 0)` | use the **weaker** side, floored at zero |

Symmetric in kinetic and energy — both share the 60% in-engine cap, so
treating them as a single resist class is mechanically motivated.

```{r echo=TRUE, message=FALSE, warning=FALSE}
no_armor_df <- no_armor_df %>%
  mutate(
    kinen_pos = pmax(kinen, 0),
    kinen_neg = pmin(kinen, 0),
    ke_floor  = pmax(pmin(kinetic, energy), 0)
  )

m0 <- lm(level ~ hardiness + fortitude + dexterity + intellect +
                 cleverness + power + kinen + nonkinen,
         data = no_armor_df)
m5 <- lm(level ~ hardiness + fortitude + dexterity + intellect +
                 cleverness + power + kinen_pos + kinen_neg + nonkinen,
         data = no_armor_df)
m6 <- lm(level ~ hardiness + fortitude + dexterity + intellect +
                 cleverness + power + kinen_pos + nonkinen,
         data = no_armor_df)
m7 <- lm(level ~ hardiness + fortitude + dexterity + intellect +
                 cleverness + power + ke_floor + nonkinen,
         data = no_armor_df)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
uber_idx <- no_armor_df$level <= 12 & no_armor_df$kinetic >= 50

cluster_sd <- function(m) {
  sd(no_armor_df$level[uber_idx] - predict(m)[uber_idx])
}

resist_compare <- data.frame(
  Model = c("M0 kinen", "M5 kinen_pos+kinen_neg", "M6 pmax(kinen,0)",
            "M7 pmax(pmin(k,e),0)"),
  R2    = c(summary(m0)$r.squared, summary(m5)$r.squared,
            summary(m6)$r.squared, summary(m7)$r.squared),
  ResidSD = c(sd(resid(m0)), sd(resid(m5)), sd(resid(m6)), sd(resid(m7))),
  AIC   = c(AIC(m0), AIC(m5), AIC(m6), AIC(m7)),
  UberCL10_SD = c(cluster_sd(m0), cluster_sd(m5), cluster_sd(m6), cluster_sd(m7))
)
knitr::kable(resist_compare, digits = c(0, 4, 3, 1, 3),
             col.names = c("Model", "R²", "Resid SD", "AIC", "Uber-CL10 SD"),
             caption = "Resist-term variants on the unarmored data")
```

M7 is unambiguously best on every metric: highest R², lowest residual SD,
lowest AIC, and crucially the tightest fit on the uber-CL10 cluster
itself (SD 0.70 vs 1.10 baseline). M5 and M6 capture the cluster's mean
but not its tightness — a creature with `kinetic = 60`, `energy = -50`
still gets `pmax(kinen, 0) = 5` of credit it apparently shouldn't.

The mechanism reading: **vulnerability on either kinetic or energy erases
all kin/eng resist credit**. Roughly half of unarmored creatures get
zero credit from this term because they have a vulnerability on one
side.

We promote M7 as the canonical unarmored form. The fitted coefficients:

```{r echo=TRUE, message=FALSE, warning=FALSE}
summary(m7)
```

Q-Q residuals:

```{r echo=FALSE, message=FALSE, warning=FALSE}
qqnorm(resid(m7)); qqline(resid(m7))
```

The Shapiro-Wilk test still rejects normality (p ≈ 3e-6), so there is
additional structure left in the unarmored residuals beyond the
vulnerability mechanism — see "Persistent Outliers" below.

# Verifying the fortitude=500 Breakpoint

The two-formula split assumes a hard discontinuity at fortitude=500. We
test it four ways: a single global linear regression with no break, a
GAM smooth, a data-driven `segmented` regression, and the proposed
hard split.

```{r echo=TRUE, message=FALSE, warning=FALSE}
df <- normalized_df %>%
  mutate(ke_floor = pmax(pmin(kinetic, energy), 0))

# Single global linear (no break)
m_global <- lm(level ~ hardiness + fortitude + dexterity + intellect +
                       cleverness + power + ke_floor + nonkinen,
               data = df)

# GAM with smooth fortitude
m_gam <- gam(level ~ s(fortitude) + hardiness + dexterity + intellect +
                     cleverness + power + ke_floor + nonkinen,
             data = df)

# Segmented regression (data-driven breakpoint)
seg_in <- lm(level ~ fortitude + hardiness + dexterity + intellect +
                     cleverness + power + ke_floor + nonkinen, data = df)
m_seg  <- segmented(seg_in, seg.Z = ~ fortitude, psi = 500)

# Two formulas, hard split at fort = 500
arm_df <- df %>% dplyr::filter(armor == 1 & fortitude >= 500)
una_df <- df %>% dplyr::filter(armor == 0)
m_arm  <- lm(level ~ hardiness + fortitude + dexterity + intellect +
                     cleverness + power + ke_floor + nonkinen,
             data = arm_df)
m_una  <- lm(level ~ hardiness + fortitude + dexterity + intellect +
                     cleverness + power + ke_floor + nonkinen,
             data = una_df)

resid_split <- c(resid(m_arm), resid(m_una))
ss_split    <- sum(resid_split^2)
n_split     <- length(resid_split)
r2_split    <- 1 - ss_split / sum((c(arm_df$level, una_df$level) -
                                   mean(c(arm_df$level, una_df$level)))^2)
sd_split    <- sd(resid_split)
# AIC for the joint two-formula model: sum of AICs (each is independent)
aic_split   <- AIC(m_arm) + AIC(m_una)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
breakpoint_compare <- data.frame(
  Model = c("Single global linear (no break)",
            "GAM smooth fortitude",
            sprintf("Segmented regression (break at fort=%.0f)",
                    m_seg$psi[1, "Est."]),
            "Two formulas, hard split at fort=500"),
  R2    = c(summary(m_global)$r.squared,
            summary(m_gam)$r.sq,
            summary(m_seg)$r.squared,
            r2_split),
  ResidSD = c(sd(resid(m_global)), sd(resid(m_gam)),
              sd(resid(m_seg)), sd_split),
  AIC   = c(AIC(m_global), AIC(m_gam), AIC(m_seg), aic_split)
)
knitr::kable(breakpoint_compare, digits = c(0, 4, 3, 1),
             col.names = c("Model", "R²", "Resid SD", "AIC"),
             caption = "Breakpoint test (level ~ stats with different fortitude treatments)")
```

The hard split at fortitude=500 beats the single global linear by ~500
AIC points and beats a flexible GAM smooth by ~50. That second margin is
the important one: it says the relationship has a true *discontinuity*
at 500, not just a slope change. Classic step function in the game code.

The sign flip across the break is iron-clad:

```{r echo=FALSE, message=FALSE, warning=FALSE}
fort_compare <- data.frame(
  Subset = c("Unarmored (fortitude < 500)", "Armored (fortitude >= 500)"),
  N      = c(nobs(m_una), nobs(m_arm)),
  Fortitude_coef = c(coef(m_una)["fortitude"], coef(m_arm)["fortitude"]),
  P_value = c(summary(m_una)$coef["fortitude", "Pr(>|t|)"],
              summary(m_arm)$coef["fortitude", "Pr(>|t|)"])
)
knitr::kable(fort_compare, digits = c(0, 0, 5, -1),
             col.names = c("Subset", "n", "Fortitude coef", "p-value"),
             caption = "Fortitude coefficient on either side of the break")
```

Both coefficients are extremely significant, and they have opposite signs.

# The Negative-Fortitude Coefficient: Artifact, Mechanism, or Bug?

The unarmored fortitude coefficient is **negative**. Taken at face value,
this would say that a creature with more fortitude has a *lower* level —
counterintuitive enough to deserve scrutiny. There are three families of
explanation:

1. **Artifact**: the coefficient is not really about fortitude — it is
   bookkeeping that compensates for a collinearity (e.g. hardiness and
   fortitude correlate 0.85). Drop fortitude or shrink it, and another
   predictor should absorb the signal with no loss of fit.
2. **Mechanism**: the game's level formula literally subtracts a
   fortitude term for unarmored creatures. Counterintuitive as gameplay,
   but mechanically possible.
3. **Bug or balance lever**: SWG ran a major creature-system rebalance
   post-launch (pre-launch pets could equip medium armor; that was
   stripped). A developer could have flipped the sign on the unarmored
   fortitude term either accidentally or as a deliberate "armor must
   pay for itself" damper.

We can rule out (1) with the data, but the data alone cannot
distinguish (2) from (3). Four tests, in order.

## Test 1: drop fortitude — does another coefficient absorb it?

Within the unarmored subset, fortitude correlates 0.85 with hardiness
and 0.84 with health — Bio-Engineers wire these up together when
crafting. If the negative-fortitude coefficient is bookkeeping for the
hardiness correlation, dropping fortitude from the model should let
hardiness pick up the signal with no loss of fit.

```{r echo=TRUE, message=FALSE, warning=FALSE}
m_una_full <- m_una
m_una_no_fort <- lm(level ~ hardiness + dexterity + intellect +
                            cleverness + power + ke_floor + nonkinen,
                    data = una_df)

shift_compare <- data.frame(
  Term = c("hardiness", "dexterity", "intellect",
           "cleverness", "power", "ke_floor", "nonkinen"),
  With_fortitude = coef(m_una_full)[c("hardiness","dexterity","intellect",
                                      "cleverness","power","ke_floor","nonkinen")],
  Without_fortitude = coef(m_una_no_fort)[c("hardiness","dexterity","intellect",
                                            "cleverness","power","ke_floor","nonkinen")]
)
shift_compare$Shift <- shift_compare$Without_fortitude - shift_compare$With_fortitude
knitr::kable(shift_compare, digits = 4, row.names = FALSE,
             caption = "Coefficient shifts when fortitude is dropped from the unarmored model")
```

The hardiness coefficient flips from +0.013 to roughly zero — exactly
what the artifact hypothesis predicts. So far, the picture is
consistent with the negative coefficient being collinearity bookkeeping.

A manual VIF (variance inflation factor) confirms the collinearity is
real:

```{r echo=TRUE, message=FALSE, warning=FALSE}
manual_vif <- function(model) {
  X <- model.matrix(model)[, -1, drop = FALSE]
  vifs <- sapply(seq_len(ncol(X)), function(j) {
    fit <- lm(X[, j] ~ X[, -j])
    1 / (1 - summary(fit)$r.squared)
  })
  setNames(vifs, colnames(X))
}
round(manual_vif(m_una_full), 2)
```

Fortitude itself is at moderate VIF (~5); intellect and cleverness are
worse. This test is *suggestive* of artifact, but not decisive — the
fact that hardiness can absorb the signal does not prove the signal is
*nothing but* hardiness.

## Test 2: constrained regression — what is the predictive cost of forcing fortitude ≥ 0?

If the negative coefficient is pure artifact, forcing fortitude ≥ 0
(equivalent to setting it to zero, since the unconstrained estimate is
on the wrong side of the constraint) should cost no predictive
performance. Hardiness will absorb the signal and the model will fit
just as well.

By KKT conditions for inequality-constrained least squares: when the
unconstrained estimate violates the constraint, the constrained
optimum sets the violating coefficient to its boundary (zero) and
re-fits the others. So the constrained regression is mathematically
identical to the drop-fortitude `lm` above; we just look at the
*predictive* metrics this time.

```{r echo=TRUE, message=FALSE, warning=FALSE}
una_df$resid_unc <- resid(m_una_full)
una_df$resid_con <- resid(m_una_no_fort)

cat(sprintf("Unconstrained M7: R^2 = %.4f, resid SD = %.3f, AIC = %.1f\n",
            summary(m_una_full)$r.squared, sd(resid(m_una_full)),
            AIC(m_una_full)))
cat(sprintf("Constrained (fort >= 0): R^2 = %.4f, resid SD = %.3f, AIC = %.1f\n",
            summary(m_una_no_fort)$r.squared, sd(resid(m_una_no_fort)),
            AIC(m_una_no_fort)))
cat(sprintf("Delta R^2 = %+.4f, delta SD = %+.3f, delta AIC = %+.1f\n",
            summary(m_una_no_fort)$r.squared - summary(m_una_full)$r.squared,
            sd(resid(m_una_no_fort)) - sd(resid(m_una_full)),
            AIC(m_una_no_fort) - AIC(m_una_full)))
```

The cost is real and substantial: ~3 percentage points of R², half a
level of residual SD, and ~140 AIC points. The artifact hypothesis
predicts ≈ zero cost — this is a clear rejection.

The cost is not uniformly distributed. Where does it land?

```{r echo=FALSE, message=FALSE, warning=FALSE}
una_df$band <- cut(una_df$fortitude,
                   breaks = c(-Inf, 0, 100, 200, 300, 400, 500),
                   labels = c("<=0","0-100","100-200","200-300","300-400","400-500"),
                   include.lowest = TRUE, right = FALSE)
band_summary <- una_df %>%
  group_by(band) %>%
  summarise(n = n(),
            sd_unc   = sd(resid_unc),
            sd_con   = sd(resid_con),
            mean_unc = mean(resid_unc),
            mean_con = mean(resid_con),
            .groups = "drop")
knitr::kable(band_summary, digits = 3,
             col.names = c("Fortitude band","n","SD (unc)","SD (con)",
                           "Mean resid (unc)","Mean resid (con)"),
             caption = "Residuals by fortitude band, unconstrained vs constrained")
```

The constrained model develops a clean U-shape across fortitude bands:
positive bias at low fortitude (under-prediction by ~0.6 levels in the
0–100 band) and negative bias at high fortitude (over-prediction by
~1.3 levels in the 400–500 band). That's the residual signature of a
*missing linear term in fortitude*. The empirical penalty was filling
exactly that hole.

The uber-CL10 cluster is the most affected:

```{r echo=TRUE, message=FALSE, warning=FALSE}
uber_idx <- una_df$level <= 12 & una_df$kinetic >= 50
cat(sprintf("Uber-CL10 cluster (n = %d):\n", sum(uber_idx)))
cat(sprintf("  Unconstrained: mean resid %+.2f, SD %.2f\n",
            mean(una_df$resid_unc[uber_idx]),
            sd(una_df$resid_unc[uber_idx])))
cat(sprintf("  Constrained:   mean resid %+.2f, SD %.2f\n",
            mean(una_df$resid_con[uber_idx]),
            sd(una_df$resid_con[uber_idx])))
```

Same spread, but a 1.5-level downward bias. The negative fortitude
coefficient was specifically pulling these creatures from "predicted
~12" down to "actually CL 10."

## Test 3: effective vs special resists

A more subtle artifact possibility: in the game, kinetic and energy
resists come from two sources — **effective** resists derived directly
from fortitude (`floor(fortitude / 10)` for unarmored) and **special**
resists from DNA samples. When both are present, special wins and
effective drops out. For creatures whose displayed kinetic/energy *is*
effective, the column M7 consumes is literally a function of fortitude
— a potential double-count that could create a spurious coefficient.

```{r echo=TRUE, message=FALSE, warning=FALSE}
unarm <- normalized_df %>% dplyr::filter(armor == 0) %>%
  mutate(
    ke_floor    = pmax(pmin(kinetic, energy), 0),
    k_effective = !is.na(kinetic.effective) & kinetic.effective != 0,
    e_effective = !is.na(energy.effective)  & energy.effective  != 0,
    k_special   = !is.na(kinetic.special)   & kinetic.special   != 0,
    e_special   = !is.na(energy.special)    & energy.special    != 0
  )

# Verify: kinetic.effective = floor(fortitude / 10) for unarmored
unarm_eff <- unarm %>% dplyr::filter(k_effective)
exact_match <- sum(unarm_eff$kinetic.effective == floor(unarm_eff$fortitude / 10))
cat(sprintf("kinetic.effective matches floor(fortitude/10): %d / %d\n",
            exact_match, nrow(unarm_eff)))

# Refit on the all-special subset (no fortitude->resist algebraic relationship)
all_special <- unarm %>%
  dplyr::filter(!k_effective & !e_effective &
                (k_special | kinetic == 0) &
                (e_special | energy == 0))
m_special <- lm(level ~ hardiness + fortitude + dexterity + intellect +
                        cleverness + power + ke_floor + nonkinen,
                data = all_special)

cat(sprintf("Full unarmored fortitude coef:  %.5f (n=%d)\n",
            coef(m_una)["fortitude"], nobs(m_una)))
cat(sprintf("All-special  fortitude coef:    %.5f (n=%d)\n",
            coef(m_special)["fortitude"], nobs(m_special)))
```

The fortitude coefficient is essentially identical on the all-special
subset, where there is *no* algebraic relationship between fortitude
and the kinetic/energy values M7 consumes. The effective-resist
double-count is ruled out as the cause.

(For completeness: zeroing out effective resists from `ke_floor` makes
the fit worse by ~40 AIC. The level formula appears to credit
effective and special resists at the same rate, using whichever value
is displayed.)

## Test 4: elastic-net under cross-validation

The strongest test: regularize the regression and let cross-validation
decide how much credibility each coefficient deserves. Lasso and
elastic-net penalize the absolute size of coefficients, so a fragile
coefficient — one that's only there to bookkeep a collinearity —
should shrink toward zero or get dropped entirely. A robust signal
survives.

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(glmnet)

predictors <- c("hardiness","fortitude","dexterity","intellect",
                "cleverness","power","ke_floor","nonkinen")
X_una <- as.matrix(una_df[, predictors])
y_una <- una_df$level

set.seed(2026)
cv_ridge <- cv.glmnet(X_una, y_una, alpha = 0.0, nfolds = 10, standardize = TRUE)
cv_enet  <- cv.glmnet(X_una, y_una, alpha = 0.5, nfolds = 10, standardize = TRUE)
cv_lasso <- cv.glmnet(X_una, y_una, alpha = 1.0, nfolds = 10, standardize = TRUE)

reg_compare <- data.frame(
  Term = c("(Intercept)", predictors),
  OLS = round(coef(m_una), 5),
  Ridge_min = round(as.vector(coef(cv_ridge, s = "lambda.min")), 5),
  ElasticNet_min = round(as.vector(coef(cv_enet, s = "lambda.min")), 5),
  Lasso_min = round(as.vector(coef(cv_lasso, s = "lambda.min")), 5)
)
knitr::kable(reg_compare, row.names = FALSE,
             caption = "Cross-validated coefficients at lambda.min, original scale")
```

Lasso and elastic-net at the cross-validated optimum *essentially
reproduce the OLS solution*, including the −0.019 fortitude
coefficient. Ridge halves both ends of the (hardiness, fortitude)
pair but keeps the negative sign on fortitude.

The lasso path is even more telling. Walking from λ=∞ down to λ=0,
predictors enter the model in order of how much signal they carry:

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5}
lasso_path <- glmnet(X_una, y_una, alpha = 1.0, standardize = TRUE)
plot(lasso_path, xvar = "lambda", label = TRUE)
title("Lasso coefficient path (unarmored M7)", line = 2.5)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
beta_path <- as.matrix(lasso_path$beta)
entry_lambda <- apply(beta_path, 1, function(row) {
  nz <- which(row != 0)
  if (length(nz) == 0) NA else lasso_path$lambda[min(nz)]
})
entry_signs <- sapply(seq_along(entry_lambda), function(i) {
  nz <- which(beta_path[i, ] != 0)
  if (length(nz) == 0) "—" else
    if (beta_path[i, min(nz)] > 0) "+" else "-"
})
entry_order <- data.frame(
  Predictor = names(entry_lambda),
  EnteredAtLambda = round(entry_lambda, 5),
  EnteredSign = entry_signs
)
entry_order <- entry_order[order(-entry_order$EnteredAtLambda), ]
knitr::kable(entry_order, row.names = FALSE,
             caption = "Lasso entry order: when each predictor first becomes nonzero")
```

`fortitude` enters at λ ≈ 0.39 with the negative sign already, and
stays negative across the entire path. Crucially, **`hardiness`
enters last** — even after fortitude. The "fragile" coefficient of the
collinear pair is hardiness, not fortitude.

## Reading the evidence

Putting the four tests together:

| Test | Predicted under "artifact" | Observed |
|---|---|---|
| Drop fortitude | No fit loss; another coef absorbs it | Hardiness flips to ~0; **R² drops 3 pts** |
| Constrained regression | No fit loss | **R² drops 3 pts; U-shape in residuals** |
| Effective resists | Coefficient explained by fort→resist channel | Coefficient identical on all-special subset |
| Elastic-net (lasso) | fortitude shrinks to zero | **fortitude stays at −0.019; hardiness enters last** |

The artifact hypothesis is rejected. The negative fortitude
coefficient is doing real, irreducible predictive work — it cannot
be reshuffled away by another coefficient and it cannot be
regularized to zero without paying a real penalty.

That leaves two live mechanisms:

- **Real game-formula term**. The unarmored CL formula literally
  subtracts a fortitude term. A "fortitude penalty" reads as
  counterintuitive gameplay, but it is mechanically possible — and
  it would explain the BE-folklore observation that pushing
  fortitude toward 499 minimized its CL contribution.
- **Bug or post-rebalance damper**. Pre-launch SWG allowed pets to
  equip medium armor; that was removed in a major creature-system
  rebalance. Either the developer flipped the sign accidentally, or
  introduced the negative term deliberately as an "armor pays for
  itself" damper to keep heavy-resist unarmored pets from being
  too tameable. Either way, the empirical coefficient is what the
  retail game ran, and that is what an emulator must reproduce.

The data does not let us choose between these two — both predict
exactly the empirical coefficient pattern we observe. But for the
purposes of reproducing the retail formula, the choice does not
matter: the coefficient is real, and the SWGEmu re-implementation
should keep it.

# Combined Fit: Both Formulas

Put the two formulas together with a hard switch at fortitude=500:

```{r echo=TRUE, message=FALSE, warning=FALSE}
source("R/creature_level_model.R")

normalized_df$predicted_level <- custom_model(normalized_df)
normalized_df$residuals       <- normalized_df$level - normalized_df$predicted_level

cat(sprintf("Combined: n=%d, R^2=%.4f, residual SD=%.3f\n",
            nrow(normalized_df),
            1 - sum(normalized_df$residuals^2) /
                sum((normalized_df$level - mean(normalized_df$level))^2),
            sd(normalized_df$residuals)))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
normalized_df <- normalized_df %>%
  mutate(armor_label = ifelse(armor == 1 & fortitude >= 500,
                              "Armored (fort >= 500)", "Unarmored"))

ggplot(normalized_df, aes(x = predicted_level, y = level, color = armor_label)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, color = "black", linetype = "dashed") +
  scale_color_manual(values = c("Armored (fort >= 500)" = "blue",
                                "Unarmored" = "red")) +
  labs(title = "Two-formula model: predicted vs actual level",
       subtitle = "M7 weaker-resist-floor for unarmored; clean linear for armored",
       x = "Predicted level", y = "Actual level", color = NULL) +
  theme_minimal()
```

# Final Formulas

**Armored (fortitude >= 500)** — clean fractions, R² ≈ 0.98:

```
level = -23
      + 0.01   * hardiness
      + 0.06   * fortitude
      + 0.005  * dexterity
      + 0.01   * intellect
      + 0.025  * cleverness
      + 0.015  * power
      + 0.10   * kinen
      + 0.08   * nonkinen
```

with `kinen = (kinetic + energy) / 2` and
`nonkinen = (blast + heat + cold + electricity + acid + stun) / 6`.

**Unarmored (fortitude < 500)** — M7 weaker-resist-floor, R² ≈ 0.95:

```
level = 7.87
      + 0.013  * hardiness
      - 0.019  * fortitude
      + 0.004  * dexterity
      + 0.010  * intellect
      + 0.025  * cleverness
      + 0.015  * power
      + 0.167  * pmax(pmin(kinetic, energy), 0)
      + 0.054  * nonkinen
```

The kinetic/energy resist term is the **weaker side, floored at zero**
— vulnerability on either kinetic or energy erases all credit for the
kin/eng resist class. Symmetric in kinetic and energy.

Combined (both formulas, n=370): R² ≈ 0.98, residual SD ≈ 1.9 levels.

# Persistent Outliers

After M7, residual non-normality is mild but still present (Shapiro-Wilk
p ≈ 3e-6 on the unarmored set). Two structural patterns survive:

1. **High-DPS under-prediction**: a handful of high-cleverness/power
   creatures (e.g. rancor CL48, falumpaset CL32, razor cat CL49)
   under-predict by 5–7 levels. Earlier scratchpad work suggested
   diminishing-returns thresholds at cleverness 200/300/400 — worth
   revisiting under M7.
2. **Near-armor-border zone**: residuals concentrate at
   fortitude 380–540, hinting that the hard `fortitude < 500` step in
   the implementation may not match the game exactly — possibly a
   transition band the two formulas don't bridge cleanly.

A 10-creature "data-quality suspects" sensitivity analysis (low-
health-for-skin entries plus the 6 effective-resist formula mismatches)
shifts coefficients by less than 0.01 each and improves combined SD
from 1.92 to 1.88. Coefficient stability under that ablation gives us
confidence the formula is not being driven by a small number of
suspect rows.

# Implementation Pseudocode

```c
int calculateCreatureLevel(const Creature* c) {
    float kinen = (c->kinetic + c->energy) / 2.0f;
    float nonkinen = (c->blast + c->heat + c->cold +
                      c->electricity + c->acid + c->stun) / 6.0f;

    float level;

    if (c->fortitude >= 500) {
        // ARMORED FORMULA
        level = -23.0f
              + 0.01f   * c->hardiness
              + 0.06f   * c->fortitude
              + 0.005f  * c->dexterity
              + 0.01f   * c->intellect
              + 0.025f  * c->cleverness
              + 0.015f  * c->power
              + 0.1f    * kinen
              + 0.08f   * nonkinen;
    } else {
        // UNARMORED FORMULA — M7 weaker-resist-floor.
        // ke_floor = max(0, min(kinetic, energy)): if either side has a
        // vulnerability, the kin/eng resist credit collapses to zero.
        float ke_floor = fmaxf(0.0f, fminf(c->kinetic, c->energy));

        level = 7.87f
              + 0.013f  * c->hardiness
              - 0.019f  * c->fortitude
              + 0.004f  * c->dexterity
              + 0.010f  * c->intellect
              + 0.025f  * c->cleverness
              + 0.015f  * c->power
              + 0.167f  * ke_floor
              + 0.054f  * nonkinen;
    }

    if (level < 1.0f)  level = 1.0f;

    return (int)(level + 0.5f);   // round to nearest
}
```

The R counterpart of this pseudocode lives in `R/creature_level_model.R`
as `custom_model()`.
