---
title: "Damage Analysis"
output: html_document
---

# Data

The GBM of Furrycat's data shows a strong relationship between damage and power.

```{r setup}
library(magrittr)
library(dplyr)
library(readr)
library(ggplot2)
library(gbm)
```

```{r}

# Some of these are obvious typos -- fix later
bad_data <- c("dta2275t", "r8tfng1v",              # weird hardiness
              "er9nq4et", "38u9raer", "6ko7k4ql",  # weird action/dex
              "6td1segp", "6f9lmdbu",              # bad speed
              "frfksnne",                          # Maybe bad to-hit but interesting...
              "kjvtv7d7",                          # has fortitude < 500 AND light armor?
              "4231uemm",                          # trash mob in piket skin (min cl 25) that throws off CL calculations
              "at8d24tc",                          # rancor skin (min cl 35) also throwing off cl calculations
              "8mgpmeev",                          # thune skin also throwing off cl
              "v22scdcg",                          # fambaa
              "ceh4m90v",                          # kimogila
              "mm8gn9sj"                           # kimogila
)
creatures <- read_csv("data/clean/furrycat/creature.csv") %>%
  filter(!creature_id %in% bad_data)
final_exp <- read_csv("data/clean/furrycat/final_experiment.csv")
combined_data <- inner_join(creatures, final_exp, by = c("final_experimentation_id" = "serial"))
normalized_df <- combined_data %>%
  mutate_at(
    vars("kinetic.x", "energy.x", "blast.x", "heat.x", "cold.x", "electricity.x", "acid.x", "stun.x"),
    ~ ifelse(. < 0, -99, .)
  ) %>%
  mutate_at(
    vars("kinetic.x", "energy.x", "blast.x", "heat.x", "cold.x", "electricity.x", "acid.x", "stun.x"),
    ~ ifelse(is.na(.), 0, .)
  )

gbm_model_low <- gbm(
  damage_low ~ hardiness + fortitude + dexterity + endurance + intellect + cleverness + courage + dependability + fierceness + power + armor.x + kinetic.x + energy.x + blast.x + heat.x + cold.x +  electricity.x + acid.x + stun.x,
  data = normalized_df,
  distribution = "gaussian",
  n.trees = 500,
  interaction.depth = 3,
  n.minobsinnode = 5,
  bag.fraction = 0.5,
  cv.folds = 10
)
gbm_model_high <- gbm(
  damage_high ~ hardiness + fortitude + dexterity + endurance + intellect + cleverness + courage + dependability + fierceness + power + armor.x + kinetic.x + energy.x + blast.x + heat.x + cold.x +  electricity.x + acid.x + stun.x,
  data = normalized_df,
  distribution = "gaussian",
  n.trees = 500,
  interaction.depth = 3,
  n.minobsinnode = 5,
  bag.fraction = 0.5,
  cv.folds = 10
)
```

```{r}
summary(gbm_model_low)
```

```{r}
summary(gbm_model_high)
```


The linear model shows high correlation and low residuals.


```{r}
model_low <- lm(damage_low ~ power, data = normalized_df)
summary(model_low)
```

```{r}
model_high <- lm(damage_high ~ power, data = normalized_df)
summary(model_high)
```

And looks like this.


```{r}
normalized_df$predicted_damage_low <- predict.lm(model_low, newdata = normalized_df)
normalized_df$predicted_damage_high <- predict.lm(model_high, newdata = normalized_df)

ggplot(normalized_df, aes(x = damage_low, y = predicted_damage_low)) +
  geom_point() +
  geom_smooth(method = lm, color = "blue") +
  ggtitle("Linear Model Predicted Damage_low vs Actual")

```

```{r}
ggplot(normalized_df, aes(x = damage_high, y = predicted_damage_high)) +
  geom_point() +
  geom_smooth(method = lm, color = "blue") +
  ggtitle("Linear Model Predicted Damage_High vs Actual")

```


# Outliers

The predictions above show some outliers, however it seems that high and low damage are flipped in the existing data set.

```{r}
residuals_low <- normalized_df$damage_low - normalized_df$predicted_damage_low
qqnorm(residuals_low)
qqline(residuals_low)
```

```{r}
residuals_high <- normalized_df$damage_high - normalized_df$predicted_damage_high
qqnorm(residuals_high)
qqline(residuals_high)
```

```{r}
standardized_residuals <- rstandard(model_low)
outlier_indices <- which(abs(standardized_residuals) > 2)
outlier_records <- normalized_df[outlier_indices, ]
outlier_records %>% select(creature_id, skin, creature_level, damage_low, predicted_damage_low, power, fierceness)
```

```{r}
standardized_residuals <- rstandard(model_high)
outlier_indices <- which(abs(standardized_residuals) > 2)
outlier_records <- normalized_df[outlier_indices, ]
outlier_records %>% select(creature_id, skin, creature_level, damage_high, predicted_damage_high, power, fortitude)
```


# Rounding and Clamping

Speed displays both rounding and clamping, as shown by the Gradient Boosting Model above.

```{r}
plot.gbm(gbm_model_low, i.var = "power")
```

```{r}
plot.gbm(gbm_model_high, i.var = "power")
```
# GBM Model Performance

It may be that at a certain threshold, high and low damage are flipped. GBM would be better equipped to predict this. Does it do any better?

```{r}
normalized_df$predicted_damage_low_gmb <- predict(gbm_model_low, newdata = normalized_df, n.trees = 500)
normalized_df$predicted_damage_high_gbm <- predict(gbm_model_high, newdata = normalized_df, n.trees = 500)
```

```{r}
ggplot(normalized_df, aes(x = damage_low, y = predicted_damage_low_gmb)) +
  geom_point() +
  geom_smooth(method = lm, color = "blue") +
  ggtitle("Linear Model Predicted Damage_low GBM vs Actual")
```

```{r}
ggplot(normalized_df, aes(x = damage_high, y = predicted_damage_high_gbm)) +
  geom_point() +
  geom_smooth(method = lm, color = "blue") +
  ggtitle("Linear Model Predicted Damage_low vs Actual")
```

GBM does substantially better -- nearly perfect -- so there is likely something else going on here.

# The Role of Fierceness and... Fortitude?

The next variable GBM predicts as important for low damage is fierceness, and for high damage it is... fortitude? When viewed as an independent variable, it shows something interesting.

```{r}
plot.gbm(gbm_model_low, i.var = "fierceness")
```

```{r}
plot.gbm(gbm_model_high, i.var = "fortitude")
```

There is some threshold at which point fierceness adds about 15 to damage_low and fortitude adds about 10 to damage_high. This appears to be between 400 and 550 for both attributes. From the graphs above it looks almost as if the slopes flip.

# Conclusion

Minimum Damage is roughly captured by the following formula, $damage_{low} \approx 14 + 0.76784 * power$.

Maximum Damage is roughly captured by the following formula, $damage_{max} \approx 21 + 0.794045 * power$.

Both attributes seem to be rounded to the nearest 5.

The exact algorithm for damage is likely complicated by rounding and clamping. Furthermore, there is evidence that some threshold at about 400 fierceness effects the minimum damage calculation; and a threshold at 500 fortitude effects the maximum damage calculation. More investigation is needed.
