---
title: "Damage Analysis"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("R/data.R")
library(mgcv)
library(broom)
library(ggplot2)
```


# Data

A GAM of Furrycat's data shows a strong relationship between damage and power.

Note that damage spread was changed near patch 14.1, so many of the examples in the Furrycat data set are not valid.

- https://archive.cuemu.com/thread/28261
- https://archive.cuemu.com/thread/28182
- https://archive.cuemu.com/thread/28257
- https://archive.cuemu.com/thread/28312

We can see this immediately from the complete data set as well.


```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyr)
normalized_df <- normalized_df %>% mutate(damage_spread = damage_high - damage_low) %>% filter(serial != "4rt3st55")

long_df <- pivot_longer(
  normalized_df,
  cols = c(damage_low, damage_high),
  names_to = "damage_type",
  values_to = "damage_value"
)
ggplot(long_df, aes(x = power, y = damage_value, color = damage_type)) +
  geom_point() +
  scale_color_manual(values = c("damage_low" = "blue", "damage_high" = "red")) +
  ggtitle("Power vs. Damage")
```


Where we see a rather striking segment starting at around $damage_{min} \approx 300$.

For the purposes of fitting the data, we will remove these pre-patch examples.

```{r echo=TRUE, message=FALSE, warning=FALSE}
post_patch_dataset <- normalized_df %>%
  filter((damage_spread > 10 & power >= 380) | power < 380)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
long_df <- pivot_longer(
  post_patch_dataset,
  cols = c(damage_low, damage_high),
  names_to = "damage_type",
  values_to = "damage_value"
)
ggplot(long_df, aes(x = power, y = damage_value, color = damage_type)) +
  geom_point() +
  scale_color_manual(values = c("damage_low" = "blue", "damage_high" = "red")) +
  ggtitle("Power vs. Damage")
```

A naive general additive model shows high influence from power.


```{r echo=TRUE, message=FALSE, warning=FALSE}
model_low <- gam(
  formula = damage_low ~
    s(hardiness) +
    s(fortitude) + 
    s(dexterity) + 
    s(endurance) +
    s(intellect) + 
    s(cleverness) + 
    s(courage) + 
    s(dependability) +
    s(power) +
    s(fierceness) +
    armor,
  family = gaussian(),
  data = post_patch_dataset
)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
summary(model_low)
```

And $damage_{high}$ is best expressed in terms of $damage_{low}$.

```{r echo=TRUE, message=FALSE, warning=FALSE}
model_high <- gam(
  formula = damage_high ~ s(damage_low),
  family = gaussian(),
  data = post_patch_dataset
)
summary(model_high)
```

A simple change detection and segmentation analysis gives the piecewise-defined regressions.

```{r echo=TRUE, message=FALSE, warning=FALSE}
model <- lm(damage_low ~ power, data = post_patch_dataset)
summary(model)
```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
library(segmented)
segmented.model <- segmented(model, seg.Z = ~power, psi=380)
summary(segmented.model)
```

Which visually fits.

```{r echo=FALSE, message=FALSE, warning=FALSE}
post_patch_dataset$predicted_damage_low <-segmented.model$fitted.values
post_patch_dataset$predicted_damage_low_residuals <- post_patch_dataset$damage_low - post_patch_dataset$predicted_damage_low

ggplot(post_patch_dataset, aes(x = power, y = damage_low)) +
  geom_point() +
  geom_line(aes(x = power, y = predicted_damage_low), color = 'red', linetype = 'dotted')

```

What about $damage_{high}$? That's even more simple, especially if expressed in terms of $damage_{low}$.

```{r echo=TRUE, message=FALSE, warning=FALSE}
model <- lm(damage_high ~ damage_low, data = post_patch_dataset)
segmented.model <- segmented(model, seg.Z = ~damage_low, psi=300)
summary(segmented.model)
```

Which also fits.

```{r echo=FALSE, message=FALSE, warning=FALSE}
post_patch_dataset$predicted_damage_high <-segmented.model$fitted.values
post_patch_dataset$predicted_damage_high_residuals <- post_patch_dataset$damage_high - post_patch_dataset$predicted_damage_high

ggplot(post_patch_dataset, aes(x = power, y = damage_high)) +
  geom_point() +
  geom_line(aes(x = power, y = predicted_damage_high), color = 'red', linetype = 'dotted')

```

# Conclusion

Damage is roughly captured by the following formulas, 

$$
 \begin{equation}
 damage_{low} \approx
   \left\{\begin{array}{lr}
       max(13 + (7/9) * power, 30), & 0 < power \le 369 \\
       300 + 0.5 * power, & power > 369
    \end{array}\right.
 \end{equation}
$$

$$
 \begin{equation}
 damage_{high} \approx
   \left\{\begin{array}{lr}
       10 + damage_{low}, & damage_{low} \le 300 \\
       2 * damage_{low} - 290, & damage_{low} > 300
    \end{array}\right.
 \end{equation}
$$ 

Furthermore, $damage_{high}$ will round to the nearest $10$, and $damage_{low}$ will (sometimes) round to the nearest $5$. There is more investigation needed into the rounding of these values; my belief is that $damage_{high}$ is based on the _non-rounded_ value of $damage_{low}$, and then the rounding occurs differently. More investigation on that behavior is needed, however I do consider some of that data of suspect quality, especially where the difference between $damage_{high}$ and $damage_{low}$ is allegedly $5$.
