---
title: "Damage Analysis"
output: html_document
---

# Data

The GBM of Furrycat's data shows a strong relationship between damage and power.


```{r message=FALSE, warning=FALSE, include=FALSE}
library(magrittr)
library(dplyr)
library(readr)
library(ggplot2)
library(gbm)
library(tidyr)

# Some of these are obvious typos -- fix later
bad_data <- c("dta2275t", "r8tfng1v",              # weird hardiness
              "er9nq4et", "38u9raer", "6ko7k4ql",  # weird action/dex
              "6td1segp", "6f9lmdbu",              # bad speed
              "frfksnne",                          # Maybe bad to-hit but interesting...
              "kjvtv7d7",                          # has fortitude < 500 AND light armor?
              "4231uemm",                          # trash mob in piket skin (min cl 25) that throws off CL calculations
              "at8d24tc",                          # rancor skin (min cl 35) also throwing off cl calculations
              "8mgpmeev",                          # thune skin also throwing off cl
              "v22scdcg",                          # fambaa
              "ceh4m90v",                          # kimogila
              "mm8gn9sj"                           # kimogila
)
creatures <- read_csv("data/clean/furrycat/creatures.csv") %>%
  filter(!serial %in% bad_data)
templates <- read_csv("data/clean/furrycat/templates.csv")
combined_data <- creatures %>% inner_join(templates, by = join_by(template_id == serial))
normalized_df <- combined_data %>%
  mutate_at(
    vars("kinetic.effective.x",
         "kinetic.special.x",
         "energy.effective.x",
         "energy.special.x",
         "blast.effective.x",
         "blast.special.x",
         "heat.effective.x",
         "heat.special.x",
         "cold.effective.x",
         "cold.special.x",
         "electricity.effective.x",
         "electricity.special.x",
         "acid.effective.x",
         "acid.special.x",
         "stun.effective.x",
         "stun.special.x"),
    ~ ifelse(is.na(.), 0, .)
  )

experiments <- read_csv("data/clean/furrycat/experiments.csv")
samples <- read_csv("data/clean/furrycat/samples.csv")
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
gbm_model_low <- gbm(
  damage_low ~     
    hardiness + fortitude +
    dexterity + endurance +
    intellect + cleverness +
    courage + dependability +
    fierceness + power +
    kinetic.special.x + kinetic.effective.x +
    energy.special.x + energy.effective.x +
    blast.special.x + blast.effective.x +
    heat.special.x + heat.effective.x +
    cold.special.x + cold.effective.x +
    electricity.special.x + electricity.effective.x +
    acid.special.x + acid.effective.x +
    stun.special.x + stun.effective.x,
  data = normalized_df,
  distribution = "gaussian",
  n.trees = 500,
  interaction.depth = 3,
  n.minobsinnode = 5,
  bag.fraction = 0.5,
  cv.folds = 10
)

gbm_model_high <- gbm(
  damage_high ~ 
    hardiness + fortitude +
    dexterity + endurance +
    intellect + cleverness +
    courage + dependability +
    fierceness + power +
    kinetic.special.x + kinetic.effective.x +
    energy.special.x + energy.effective.x +
    blast.special.x + blast.effective.x +
    heat.special.x + heat.effective.x +
    cold.special.x + cold.effective.x +
    electricity.special.x + electricity.effective.x +
    acid.special.x + acid.effective.x +
    stun.special.x + stun.effective.x,
  data = normalized_df,
  distribution = "gaussian",
  n.trees = 500,
  interaction.depth = 3,
  n.minobsinnode = 5,
  bag.fraction = 0.5,
  cv.folds = 10
)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
summary(gbm_model_low, plotit = FALSE)
```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
summary(gbm_model_high, plotit = FALSE)
```


And GBM shows a linear relationships.

```{r echo=TRUE, message=FALSE, warning=FALSE}
plot.gbm(gbm_model_low, i.var = "power", n.trees = 500)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
plot.gbm(gbm_model_high, i.var = "power", n.trees = 500)
```

The linear model shows high correlation and low residuals.

```{r echo=FALSE, message=FALSE, warning=FALSE}
long_df <- pivot_longer(
  normalized_df,
  cols = c(damage_low, damage_high),
  names_to = "damage_type",
  values_to = "damage_value"
)

ggplot(long_df, aes(x = power, y = damage_value, color = damage_type)) +
  geom_point() +
  scale_color_manual(values = c("damage_low" = "blue", "damage_high" = "red")) +
  ggtitle("Power vs. Damage")
```

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
model_low <- lm(damage_low ~ power, data = normalized_df)
summary(model_low)
```
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
model_high <- lm(damage_high ~ power, data = normalized_df)
summary(model_high)
```

And looks like this.


```{r echo=FALSE, message=FALSE, warning=FALSE}
normalized_df$predicted_damage_low <- predict.lm(model_low, newdata = normalized_df)

ggplot(normalized_df, aes(x = damage_low, y = predicted_damage_low)) +
  geom_point() +
  geom_smooth(method = lm, color = "blue") +
  ggtitle("Linear Model Predicted Damage Low vs Actual")

```



```{r echo=FALSE, message=FALSE, warning=FALSE}
normalized_df$predicted_damage_high <- predict.lm(model_high, newdata = normalized_df)

ggplot(normalized_df, aes(x = damage_high, y = predicted_damage_high)) +
  geom_point() +
  geom_smooth(method = lm, color = "blue") +
  ggtitle("Linear Model Predicted Damage High vs Actual")

```

# Outliers

The outlier data curiously follow a pattern. This could be due to something systematic, or perhaps an update to the game caused damage to be calculated differently. This happens when power is around 300, as can be seen above.

Curiously, in the data set above 300 power, we see power have less influence in both damage attributes; however, power is still the primary influence.

# Conclusion

Damage is roughly captured by the following formulas, 

$$
damage_{low} \approx 14.44471 + 0.76784 * power
$$
$$
damage_{high} \approx 20.973211 + 0.794045 * power
$$ 

However, more investigation is needed.
