---
title: "Health Analysis"
output: html_document
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("R/data.R")
library(mgcv)
library(broom)
library(ggplot2)
```

# Data

A GAM of Furrycat's data shows a strong relationship between health and both hardiness and dexterity.

```{r echo=TRUE, message=FALSE, warning=FALSE}
model <- gam(
  formula = health ~
    s(hardiness) +
    s(fortitude) + 
    s(dexterity) + 
    s(endurance) +
    s(intellect) + 
    s(cleverness) + 
    s(courage) + 
    s(dependability) +
    s(power) +
    s(fierceness) +
    armor,
  family = gaussian(),
  data = normalized_df
)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
summary(model)
```

And GAM shows a linear relationships.

```{r echo=TRUE, message=FALSE, warning=FALSE}
plot(model, select = 1)
plot(model, select = 3)
```


The linear model shows high correlation and low residuals.


```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
model <- lm(health ~ hardiness + dexterity, data = normalized_df)
summary(model)
```

And looks like this.

```{r echo=FALSE, message=FALSE, warning=FALSE}
normalized_df$predicted_health <- predict.lm(model, newdata = normalized_df)

ggplot(normalized_df, aes(x = predicted_health, y = health)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  ggtitle("Linear Model Predicted Health vs Actual")

```

The residuals suggest that some of the error may be due to rounding.

```{r echo=FALSE, message=FALSE, warning=FALSE}
res <- resid(model)
qqnorm(res)
qqline(res)
```


# Conclusion

Health is roughly captured by the following formula, $health \approx 42 + 15 * hardiness + 3 * dexterity$. The final health is rounded.

