---
title: "Speed Analysis"
output: html_document
---

# Data

The GBM of Furrycat's data shows a strong relationship between speed and courage.

```{r setup}
library(magrittr)
library(dplyr)
library(readr)
library(ggplot2)
library(gbm)
```

```{r}

# Some of these are obvious typos -- fix later
bad_data <- c("dta2275t", "r8tfng1v",              # weird hardiness
              "er9nq4et", "38u9raer", "6ko7k4ql",  # weird action/dex
              "6td1segp", "6f9lmdbu",              # bad speed
              "frfksnne",                          # Maybe bad to-hit but interesting...
              "kjvtv7d7",                          # has fortitude < 500 AND light armor?
              "4231uemm",                          # trash mob in piket skin (min cl 25) that throws off CL calculations
              "at8d24tc",                          # rancor skin (min cl 35) also throwing off cl calculations
              "8mgpmeev",                          # thune skin also throwing off cl
              "v22scdcg",                          # fambaa
              "ceh4m90v",                          # kimogila
              "mm8gn9sj"                           # kimogila
)
creatures <- read_csv("data/clean/furrycat/creature.csv") %>%
  filter(!creature_id %in% bad_data)
final_exp <- read_csv("data/clean/furrycat/final_experiment.csv")
combined_data <- inner_join(creatures, final_exp, by = c("final_experimentation_id" = "serial"))
normalized_df <- combined_data %>%
  mutate_at(
    vars("kinetic.x", "energy.x", "blast.x", "heat.x", "cold.x", "electricity.x", "acid.x", "stun.x"),
    ~ ifelse(. < 0, -99, .)
  ) %>%
  mutate_at(
    vars("kinetic.x", "energy.x", "blast.x", "heat.x", "cold.x", "electricity.x", "acid.x", "stun.x"),
    ~ ifelse(is.na(.), 0, .)
  )

gbm_model <- gbm(
  speed ~ hardiness + fortitude + dexterity + endurance + intellect + cleverness + courage + dependability + fierceness + power + armor.x + kinetic.x + energy.x + blast.x + heat.x + cold.x +  electricity.x + acid.x + stun.x,
  data = normalized_df,
  distribution = "gaussian",
  n.trees = 500,
  interaction.depth = 3,
  n.minobsinnode = 5,
  bag.fraction = 0.5,
  cv.folds = 10
)
```

```{r}
summary(gbm_model)
```

The linear model shows high correlation and low residuals.


```{r}
model <- lm(speed ~ courage, data = normalized_df)
summary(model)
```

And looks like this.


```{r}
normalized_df$predicted_speed <- predict.lm(model, newdata = normalized_df)

ggplot(normalized_df, aes(x = speed, y = predicted_speed)) +
  geom_point() +
  geom_smooth(method = lm, color = "blue") +
  ggtitle("Linear Model Predicted Speed vs Actual")

```

# Outliers

The predictions above show some interesting outliers.

```{r}
residuals <- normalized_df$speed - normalized_df$predicted_speed
qqnorm(residuals)
qqline(residuals)
```
```{r}
standardized_residuals <- rstandard(model)
outlier_indices <- which(abs(standardized_residuals) > 2)
outlier_records <- normalized_df[outlier_indices, ]
outlier_records %>% select(creature_id, skin, creature_level, speed, predicted_speed, courage)
```

# Rounding and Clamping

Speed displays both rounding and clamping, as shown by the Gradient Boosting Model above.

```{r}
plot.gbm(gbm_model, i.var = "courage")
```


# Conclusion

Speed is roughly captured by the following formula, $speed \approx 2.5 - courage / 1000$.

The exact algorithm for speed is likely complicated by rounding and clamping. Courage below 100 and above 500 do not appear to have an effect on speed. The reason for the outliers is unknown at this time, but at least one may have been due to capping speed, as courage far exceeded 500; the others are not known at this time.
