---
title: "To-Hit Analysis"
output: html_document
---

# Data

The GBM of Furrycat's data shows a strong relationship between to-hit and cleverness.

```{r setup}
library(magrittr)
library(dplyr)
library(readr)
library(ggplot2)
library(gbm)
```

```{r}

# Some of these are obvious typos -- fix later
bad_data <- c("dta2275t", "r8tfng1v",              # weird hardiness
              "er9nq4et", "38u9raer", "6ko7k4ql",  # weird action/dex
              "6td1segp", "6f9lmdbu",              # bad speed
              "frfksnne",                          # Maybe bad to-hit but interesting...
              "kjvtv7d7",                          # has fortitude < 500 AND light armor?
              "4231uemm",                          # trash mob in piket skin (min cl 25) that throws off CL calculations
              "at8d24tc",                          # rancor skin (min cl 35) also throwing off cl calculations
              "8mgpmeev",                          # thune skin also throwing off cl
              "v22scdcg",                          # fambaa
              "ceh4m90v",                          # kimogila
              "mm8gn9sj"                           # kimogila
)
creatures <- read_csv("data/clean/furrycat/creature.csv") %>%
  filter(!creature_id %in% bad_data)
final_exp <- read_csv("data/clean/furrycat/final_experiment.csv")
combined_data <- inner_join(creatures, final_exp, by = c("final_experimentation_id" = "serial"))
normalized_df <- combined_data %>%
  mutate_at(
    vars("kinetic.x", "energy.x", "blast.x", "heat.x", "cold.x", "electricity.x", "acid.x", "stun.x"),
    ~ ifelse(. < 0, -99, .)
  ) %>%
  mutate_at(
    vars("kinetic.x", "energy.x", "blast.x", "heat.x", "cold.x", "electricity.x", "acid.x", "stun.x"),
    ~ ifelse(is.na(.), 0, .)
  )

gbm_model <- gbm(
  to_hit ~ hardiness + fortitude + dexterity + endurance + intellect + cleverness + courage + dependability + fierceness + power + armor.x + kinetic.x + energy.x + blast.x + heat.x + cold.x +  electricity.x + acid.x + stun.x,
  data = normalized_df,
  distribution = "gaussian",
  n.trees = 500,
  interaction.depth = 3,
  n.minobsinnode = 5,
  bag.fraction = 0.5,
  cv.folds = 10
)
```

```{r}
summary(gbm_model)
```

The linear model shows high correlation and low residuals.


```{r}
model <- lm(to_hit ~ cleverness, data = normalized_df)
summary(model)
```

And looks like this.


```{r}
normalized_df$predicted_to_hit <- predict.lm(model, newdata = normalized_df)

ggplot(normalized_df, aes(x = to_hit, y = predicted_to_hit)) +
  geom_point() +
  geom_smooth(method = lm, color = "blue") +
  ggtitle("Linear Model Predicted To-Hit vs Actual")

```

# Outliers

The predictions above show some interesting outliers.

```{r}
residuals <- normalized_df$to_hit - normalized_df$predicted_to_hit
qqnorm(residuals)
qqline(residuals)
```
```{r}
standardized_residuals <- rstandard(model)
outlier_indices <- which(abs(standardized_residuals) > 2)
outlier_records <- normalized_df[outlier_indices, ]
outlier_records %>% select(creature_id, skin, creature_level, to_hit, predicted_to_hit, cleverness)
```

# Rounding and Clamping

Speed displays both rounding and clamping, as shown by the Gradient Boosting Model above.

```{r}
plot.gbm(gbm_model, i.var = "cleverness")
```




# Conclusion

To-Hit is roughly captured by the following formula, $to_hit \approx 1.95 + 0.0006458 * cleverness$. GBM shows clear rounding happening, so discrepancies may be due to that. Also, GBM shows that cleverness has no effect on to-hit after 500 or so.
